### 2020-04-17 doc/gen_fcn_notes.tex
Add work on writeup of gen func solutions to mRNA dists

\documentclass[12pt]{article}%{amsart}
\usepackage[top = 1.0in, bottom = 1.0in, left = 1.0in, right = 1.0in]{geometry}
% See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}               % ... or a4paper or a5paper or ...
%\geometry{landscape}             % Activate for for rotated page geometry
\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
%\usepackage{epstopdf}
% \usepackage{caption}
\usepackage{amsmath}
% \usepackage{longtable}
% \usepackage{tabu}
% \usepackage{accents} %to get undertildes for vec & mat
\usepackage{color} % for colored text
% \usepackage[normalem]{ulem} % for editing: striking out text using 'sout' command

% \newcommand\mytablefigwidth{0.35\textwidth}

% allows to use .ai files directly w/o resaving as pdf
\DeclareGraphicsRule{.ai}{pdf}{.ai}{}

% Handy math macros!
\newcommand{\vect}[1]{\vec{#1}}
\newcommand{\matr}[1]{\mathbf{#1}}
\newcommand{\rate}[3]{{#1}_{#2}^{#3}}
\newcommand{\mmnote}[1]{\textcolor{cyan}{(MM:~#1)}}

% derivative macros. these are sneaky. the [{}] is an empty default 1st arg
% usage: provide 1 arg: \deriv{x} gives d/dx
% provide 2 args like so: \deriv[f]{x} gives df/dx
\newcommand{\deriv}[2][{}]{\frac{d #1}{d #2}}
\newcommand{\pderiv}[2][{}]{\frac{\partial #1}{\partial #2}}
\newcommand{\psecderiv}[2][{}]{\frac{\partial{^2} #1}{\partial #2{^2}}}

%%%%% Referencing macros %%%%%
\newcommand{\fref}[1]{Figure~\ref{#1}}
\newcommand{\tref}[1]{Table~\ref{#1}}
\newcommand{\eref}[1]{Eq.~(\ref{#1})}
\newcommand{\erngref}[2]{Eq.~(\ref{#1}-\ref{#2})} % Equations
% \newcommand{\test}[1][]{%
% \ifthenelse{\isempty{#1}{}}{omitted}{#1}%
% }

%%%%%
\begin{document}

\title{Generating function solution for bursty promoters}

\maketitle

\section{Constitutive promoter with bursts}

\subsection{From master equation to generating function}
Before tackling mRNA and repression together, let us
just consider mRNA \mmnote{replicating Charlotte's notes; she says it's worked out
in~\cite{Paulsson2000}, but I don't see it, at least not in notation I can
comprehend}. The master equation of interest is
\begin{align}
\deriv{t}p(m,t) = (m+1)\gamma p(m+1,t) - m\gamma p(m,t) - r p(m,t)
        + r \sum_{m^\prime=0}^m G_{m-m^\prime}(\theta) p(m^\prime,t),
\label{eq:1state_unreg_003}
\end{align}
where $G_{k}(\theta)$ is the geometric distribution defined as
\begin{align}
G_{k}(\theta) = \theta(1 - \theta)^k, \, k\in\{0,1,2,\dots\}.
\end{align}
With this convention, the mean burst size $b = (1-\theta)/\theta$.
$\gamma$ and $r$ are mRNA degradation rates and transcription burst
initiation rates, resp. The last term represents all ways the system could
end up with $m$ mRNAs, having started the burst with $m^\prime$. Define
$\lambda = r/\gamma$ and nondimensionalize time by $\gamma$, giving
\begin{align}
\deriv{t}p(m,t) = (m+1)p(m+1,t) - m p(m,t) - \lambda p(m,t)
        + \lambda \sum_{m^\prime=0}^m G_{m-m^\prime}(\theta) p(m^\prime,t),
\end{align}
The probability generating function is defined as
\begin{align}
F(z,t) = \sum_{m=0}^\infty z^m p(m,t).
\end{align}
Multiply both sides of the CME by $z^m$ and sum over all $m$ to get
\begin{align}
\pderiv[F]{t} = \sum_{m=0}^\infty z^m
\left(
(m+1)p(m+1,t) - m p(m,t) - \lambda p(m,t)
+ \lambda \sum_{m^\prime=0}^m G_{m-m^\prime}(\theta) p(m^\prime,t)
\right).
\end{align}
The first two terms can be rewritten as derivatives of the
generating function, i.e.,
\begin{align}
\sum_{m=0}^\infty z^m (m+1)p(m+1,t)
=
\pderiv{z} \sum_{m=0}^\infty z^{m+1} p(m+1,t)
=
\pderiv[F]{z}
\end{align}
and
\begin{align}
\sum_{m=0}^\infty z^m mp(m,t)
=
z\pderiv{z} \sum_{m=0}^\infty z^m p(m,t)
=
z\pderiv[F]{z}.
\end{align}
The final term is the most trouble. The trick is to reverse
the default order of the sums, as in
\begin{align}
&\lambda \sum_{m=0}^\infty \sum_{m^\prime=0}^m
        z^m \theta(1 - \theta)^{m-m^\prime} p(m^\prime,t)
\\=
&\lambda\theta \sum_{m^\prime=0}^\infty z^{m^\prime} p(m^\prime,t)
        \sum_{m=m^\prime}^\infty
        z^{m-m^\prime} (1 - \theta)^{m-m^\prime}
\\=
&\frac{\lambda\theta F}{1 - (1 - \theta)z}.
\end{align}
This works because after reordering the sums, the sum on $m$ is just
a geometric series. Its result is independent of $m^\prime$, so the sum on
$m^\prime$ reduces to the definition of the generating function.
Putting these results together, the PDE for the generation function is
\begin{align}
\pderiv[F]{t} = \pderiv[F]{z} - z\pderiv[F]{z} - \lambda F
        + \frac{\lambda\theta F}{1-z(1-\theta)}.
\end{align}
Changing variables to $\xi=1-\theta$ and simplifying gives
\begin{align}
\pderiv[F]{t} + (z - 1) \pderiv[F]{z} = \frac{(z-1)\xi}{1-z\xi}\lambda F.
\label{eq:1state_unreg_015}
\end{align}
\subsection{Stead-state}
At steady-state, the PDE reduces to the ODE
\begin{align}
\deriv[F]{z} = \frac{\xi}{1-z\xi}\lambda F,
\end{align}
which we can integrate as
\begin{align}
\int \frac{dF}{F} = \int \frac{\lambda\xi dz}{1-\xi z}.
\end{align}
The initial conditions for generating functions can be subtle and confusing.
The key fact follows from the definition
$F(z,t) = \sum_m z^m p(m,t)$,
normalization requires that
$F(z=1^-,t) = \sum_m p(m,t) = 1$.\footnote{
Sometimes the generating function may be undefined \textit{at} $z=1$ but the
limit still holds. Also many authors change variables from $z$ to other
things, so one must keep track of how the normalization condition transforms.
}
Doing the integrals (and producing constant $c$) gives
\begin{align}
\ln F &= -\lambda \ln(1-\xi z) + c
\\
F &= \frac{c}{(1-\xi z)^\lambda}.
\end{align}
Only one choice for $c$ can satisfy initial conditions, producing
\begin{align}
F(z) = \left(\frac{1-\xi}{1-\xi z}\right)^\lambda
        = \left(\frac{\theta}{1 - z(1-\theta)}\right)^\lambda,
\end{align}
which is exactly the negative binomial's generating function, as expected.

\section{Adding repression}
\subsection{Deriving the generating function for mRNA distribution}
Let us move from a one-state promoter to a two-state promoter, where one state
has repressor bound and the other produces transcriptional bursts as above.
Although now we have an equation for each promoter state, otherwise
the master equation reads similarly to the one-state case, except with
additional terms corresponding to transitions between promoter states, namely
\begin{align}
\deriv{t}p_R(m,t) =& k_R^+ p_A(m,t) - k_R^- p_R(m,t)
        + (m+1)\gamma p_R(m+1,t) - m\gamma p_R(m,t)
\\
\begin{split}
\deriv{t}p_A(m,t) =& - k_R^+ p_A(m,t) + k_R^- p_R(m,t)
        + (m+1)\gamma p_A(m+1,t) - m\gamma p_A(m,t) 
\\
&- k_i p_A(m,t) + k_i \sum_{m^\prime=0}^m \theta(1-\theta)^{m-m^\prime} p_A(m^\prime,t),
\end{split}
\end{align}
where $p_R(m,t)$ is the probability of the system having $m$ mRNA copies
and having repressor bound to the promoter at time $t$, and $p_A$ is an
analogous probability to find the promoter without repressor bound.
$k_R+$ and $k_R^-$ are, respectively, the rates at which repressors
bind and unbind to and from the promoter, and
$\gamma$ is the mRNA degradation rate.
$k_i$ is the rate at which bursts initiate, and as before,
the geometric distribution of burst sizes has mean $b=(1-\theta)/\theta$.\footnote{
$\theta$ can be viewed as a Bernoulli trial probability: after each transcript
in a burst is produced, $\theta$ is the probability that the burst
terminates and $(1-\theta)$ is the probability that the burst continues.
}

Interestingly, it turns out that this problem maps exactly onto the
three-stage promoter model considered by
Shahrezaei and Swain in~\cite{Shahrezaei2008}, with relabelings.
Their approximate solution for protein distributions amounts to
the same approximation we make here in regarding the duration
of mRNA synthesis bursts as instantaneous, so their solution
for protein distributions also solves our problem of mRNA distributions.
Let us examine the analogy more closely.
They consider a two-state promoter, as we do here, but they model
mRNA as being produced one at a time and degraded, with rates
$v_0$ and $d_0$.
Then they model translation as occuring with rate $v_1$,
and protein degradation with rate $d_1$.
Now consider the limit where $v_1, d_0\rightarrow\infty$ with
their ratio $v_1/d_0$ held constant.
$v_1/d_0$ resembles the average burst size of translation from a single
mRNA: these are the rates of two Poisson processes that compete over a
transcript, which matches the story of geometrically distributed burst sizes.
So after taking this limit, on timescales slow compared to $v_1$ and $d_0$,
it appears that transcription events fire at rate $v_0$ and produce a
geometrically distributed burst of translation of mean size $v_1/d_0$,
which intuitively matches the story we have told above for mRNA
with variables relabeled.

To verify this intuitively conjectured mapping between our problem and
the solution in~\cite{Shahrezaei2008}, we continue with a careful
solution for the mRNA distribution using probability generating functions,
following the ideas sketched in~\cite{Shahrezaei2008}.
It is natural to nondimensionalize rates in the problem by $\gamma$,
or equivalently, this amounts to measuring time in units of $\gamma^{-1}$.
We are also only interested in steady state, so we set the time
derivatives to zero, giving
\begin{align}
0 =& k_R^+ p_A(m) - k_R^- p_R(m) + (m+1) p_R(m+1) - m p_R(m)
\\
\begin{split}
0 =& - k_R^+ p_A(m) + k_R^- p_R(m) + (m+1) p_A(m+1) - m p_A(m) 
\\
&- k_i p_A(m) + k_i \sum_{m^\prime=0}^m \theta(1-\theta)^{m-m^\prime} p_A(m^\prime),
\end{split}
\end{align}

The probability generating function is defined as before in the
constitutive case, except now we must introduce a generating function
for each promoter state,
\begin{align}
f_A(z) = \sum_{m=0}^\infty z^m p_A(m),
\;
f_R(z) = \sum_{m=0}^\infty z^m p_R(m).
\end{align}
Our real objective is the generating function $f(z)$ that generates
the mRNA distribution $p(m)$, independent of what state the promoter is in.
But since $p(m) = p_A(m) + p_R(m)$, it follows too that $f(z) = f_A(z) + f_R(z)$.

So as before we multiply both equations by $z^m$ and sum over all $m$.
Each individual term transforms exactly as did an analogous term in
the constitutive case, so the coupled ODEs for the generating functions read
\begin{align}
0 =& k_R^+ f_A(z) - k_R^- f_R(z) + \pderiv{z} f_R(z) - z \pderiv{z} f_R(z)
\\
\begin{split}
0 =&  - k_R^+ f_A(z) + k_R^- f_R(z) + \pderiv{z} f_A(z) - z \pderiv{z} f_A(z)
\\
&- k_i f_A(z) + k_i \frac{\theta}{1-z(1-\theta)} f_A(z),
\end{split}
\end{align}
and after changing variables $\xi = 1 - \theta$ as before
and rearranging, we have
\begin{align}
0 &= k_R^+ f_A(z) - k_R^- f_R(z) + (1-z) \pderiv{z} f_R(z)
\\
0 &=  - k_R^+ f_A(z) + k_R^- f_R(z) + (1 - z) \pderiv{z} f_A(z)
+ k_i \frac{(z-1)\xi}{1-z\xi} f_A(z),
\end{align}
We can transform this problem from two coupled first-order ODEs to a
single second-order ODE by solving for $f_A$ in the first and plugging
into the second, giving
\begin{align}
\begin{split}
0 = (1&-z) \pderiv[f_R]{z}
+ \frac{1-z}{k_R^+}
        \left(k_R^- \pderiv[f_R]{z} + \pderiv[f_R]{z} +(z-1) \psecderiv[f_R]{z}\right)
\\
&+ \frac{k_i}{k_R^+} \frac{(z-1)\xi}{1-z\xi}
        \left(k_R^- f_R + (z-1) \pderiv[f_R]{z}\right),
\end{split}
\end{align}
where, to reduce notational clutter, we have dropped the explicit $z$
dependence of $f_A$ and $f_R$. Simplifying we have
\begin{align}
0 = \psecderiv[f_R]{z}
        - \left(\frac{k_i\xi}{1-z\xi}
                + \frac{1 + k_R^- + k_R^+}{1-z}
        \right)\pderiv[f_R]{z}
        + \frac{k_i k_R^- \xi}{(1-z\xi)(1-z)}f_R.
\end{align}
This can be recognized as the hypergeometric differential equation,
with singularities at $z=1$, $z=\xi^{-1}$, and $z=\infty$. The latter
can be verified by a change of variables from $z$ to $x=1/z$,
being careful with the chain rule, and noting that $z=\infty$ is a
singular point if and only if $x=1/z=0$ is a singular point.

The standard form of the hypergeometric differential equation has its
singularities at 0, 1, and $\infty$, so to take advantage of the
standard form solutions to this ODE, we first need to transform
variables to put it into a standard form.
However, this is subtle. While any such transformation should work in
principle, the solutions are expressed most simply in the neighborhood
of $z=0$, but the normalization condition that we need to enforce
corresponds to $z=1$. The easiest path, therefore, is to find a change
of variables that maps
1 to 0, $\infty$ to $\infty$, and $\xi^{-1}$ to 1.
This is most intuitively done in two steps.

First map the $z=1$ singularity to 0 by the change of variables $v=z-1$, giving
\begin{align}
0 = \psecderiv[f_R]{v}
        + \left(\frac{k_i\xi}{(1+v)\xi - 1}
                + \frac{1 + k_R^- + k_R^+}{v}
        \right)\pderiv[f_R]{v}
        + \frac{k_i k_R^- \xi}{((1+v)\xi - 1)v}f_R.
\end{align}
Now two singularities are at $v=0$ and $v=\infty$.
The third is determined by $(1+v)\xi -1 = 0$, or $v=\xi^{-1} - 1$.
We want another variable change that maps this third singularity to 1
(without moving 0 or infinity). Changing variables again to
$w=\frac{v}{\xi^{-1} - 1} = \frac{\xi}{1-\xi} v$ fits the bill.
In other words, the combined change of variables
\begin{align}
w = \frac{\xi}{1-\xi} (z-1)
\end{align}
maps $z = \{1, \xi^{-1}, \infty\}$ to $w =\{0, 1, \infty\}$ as desired.
Plugging in, being mindful of the chain rule and noting
$(1 + v)\xi - 1 = (1 - \xi)(w - 1)$ gives
\begin{align}
0 = \left(\frac{\xi}{1-\xi}\right)^2 \psecderiv[f_R]{w}
+ \left(
        \frac{\xi k_i}{(1-\xi)(w-1)} + \frac{\xi(1 + k_R^- + k_R^+)}{(1-\xi)w}
\right) \frac{\xi}{1-\xi} \pderiv[f_R]{w}
+ \frac{k_i k_R^- \xi^2}{(1-\xi)^2 w(w-1)}f_R.
\end{align}
This is close to the standard form of the hypergeometric differential
equation, and some cancellation and rearrangement gives
\begin{align}
0 = w(w-1)\psecderiv[f_R]{w}
+ \left(k_i w + (1 + k_R^- + k_R^+)(w-1)\right) \pderiv[f_R]{w}
+ k_i k_R^- f_R.
\end{align}
and a little more algebra produces
\begin{align}
0 = w(1-w)\psecderiv[f_R]{w}
+ \left(1 + k_R^- + k_R^+
        - (1 + k_i + k_R^- + k_R^+)w
\right) \pderiv[f_R]{w}
- k_i k_R^- f_R,
\end{align}
which is the standard form. From this we can read off the solution
in terms of hypergeometric functions ${_2F_1}$ from
any standard source, e.g.~\cite{Abramovitz}, and identify the
conventional parameters in terms of our model parameters.
We want the general solution in the neighborhood of $w=0$ ($z=1$),
which for a homogeneous linear second order ODE must be a sum of two
linearly independent solutions.
More precisely,
\begin{align}
f_R(w) = C_1 {_2F_1}(\alpha, \beta, \gamma; w)
+ C_2 w^{1-\gamma}{_2F_1}(1+\alpha-\gamma, 1+\beta-\gamma, 2-\gamma; w)
\end{align}
with parameters determined by
\begin{align}
\begin{split}
\alpha\beta &= k_i k_R^-
\\
1+\alpha+\beta &= 1+k_i+k_R^-+k_R^+
\\
\gamma &= 1 + k_R^- + k_R^+
\end{split}
\end{align}
and constants $C_1$ and $C_2$ to be set by boundary conditions.
Solving for $\alpha$ and $\beta$, we find\footnote{
Note that $\alpha$ and $\beta$ are interchangeable in the definition of
${_2F_1}$ and differ only in the sign preceeding the radical.
}
\begin{align}
\begin{split}
\alpha &= \frac{1}{2}
\left(k_i+k_R^-+k_R^+ + \sqrt{(k_i+k_R^-+k_R^+)^2 - 4k_i k_R^-}\right)
\\
\beta &= \frac{1}{2}
\left(k_i+k_R^-+k_R^+ - \sqrt{(k_i+k_R^-+k_R^+)^2 - 4k_i k_R^-}\right)
\\
\gamma &= 1 + k_R^- + k_R^+.
\end{split}
\end{align}
Since the normalization condition requires that $f_R$ be finite at $w=0$,
we can immediately set $C_2=0$ to discard the second solution.
This is because all the rate constants are strictly positive,
so $\gamma>1$ and therefore $w^{1-\gamma}$ blows up as $w\rightarrow0$.
Now that we have $f_R$, we would like to find the generating function
for the mRNA distribution, $f(z) = f_A(z) + f_R(z)$.
We can recover $f_A$ from our solution for $f_R$, namely
\begin{align}
f_A(z) = \frac{1}{k_R^+}\left(k_R^- f_R(z) + (z-1) \pderiv[f_R]{z}\right)
\end{align}
or
\begin{align}
f_A(w) = \frac{1}{k_R^+}\left(k_R^- f_R(w) + w \pderiv[f_R]{w}\right),
\end{align}
where in the second line we transformed our original relation between
$f_R$ and $f_A$ to our new, more convenient, variable $w$.
Plugging our solution for $f_R(w) = C_1{_2F_1}(\alpha, \beta, \gamma; w)$
into $f_A$, we will require the differentiation rule for ${_2F_1}$,
which tells us
\begin{align}
\pderiv[f_R]{w} = C_1\frac{\alpha\beta}{\gamma}
                {_2F_1}(\alpha+1, \beta+1, \gamma+1; w),
\end{align}
from which it follows that
\begin{align}
f_A(w) = \frac{C_1}{k_R^+}
\left(
k_R^- {_2F_1}(\alpha, \beta, \gamma; w)
+ w\frac{\alpha\beta}{\gamma} {_2F_1}(\alpha+1, \beta+1, \gamma+1; w)
\right)
\end{align}
and therefore
\begin{align}
f(w) = C_1\left(1 + \frac{k_R^-}{k_R^+}\right)
        {_2F_1}(\alpha, \beta, \gamma; w)
+ w \frac{C_1}{k_R^+} \frac{\alpha\beta}{\gamma}
        {_2F_1}(\alpha+1, \beta+1, \gamma+1; w).
\end{align}
To proceed, we need one of the (many) useful identities known for
hypergeometric functions, in particular
\begin{align}
w\frac{\alpha\beta}{\gamma} {_2F_1}(\alpha+1, \beta+1, \gamma+1; w)
=
(\gamma-1)\left(
{_2F_1}(\alpha, \beta, \gamma-1; w) - {_2F_1}(\alpha, \beta, \gamma; w)
\right).
\end{align}
Substituting this for the second term in $f(w)$, we find
\begin{align}
f(w) = \frac{C_1}{k_R^+}
\left[
        \left(k_R^+ + k_R^-\right)
        {_2F_1}(\alpha, \beta, \gamma; w)
+ (\gamma-1)\left(
        {_2F_1}(\alpha, \beta, \gamma-1; w) - {_2F_1}(\alpha, \beta, \gamma; w)
        \right)
\right],
\end{align}
and since $\gamma-1 = k_R^+ + k_R^-$, the first and third terms cancel,
leaving only
\begin{align}
f(w) = C_1\frac{k_R^+ + k_R^-}{k_R^+} {_2F_1}(\alpha, \beta, \gamma-1; w).
\end{align}
Now we enforce normalization, demanding $f(w=0) = f(z=1) = 1$.
${_2F_1}(\alpha, \beta, \gamma-1; 0) = 1$, so we must have
$C_1 = k_R^+ / (k_R^+ + k_R^-)$ and consequently
\begin{align}
f(w) =  {_2F_1}(\alpha, \beta, k_R^+ + k_R^-; w).
\end{align}
Recalling that the mean burst size $b = (1-\theta)/\theta = \xi/(1-\xi)$
and $w = \frac{\xi}{1-\xi} (z-1) = b (z-1)$,
we can transform back to the original variable $z$ to find the tidy result
\begin{align}
f(z) =  {_2F_1}(\alpha, \beta, k_R^+ + k_R^-; b(z-1)),
\end{align}
with $\alpha$ and $\beta$ given above by
\begin{align}
\begin{split}
\alpha &= \frac{1}{2}
\left(k_i+k_R^-+k_R^+ + \sqrt{(k_i+k_R^-+k_R^+)^2 - 4k_i k_R^-}\right)
\\
\beta &= \frac{1}{2}
\left(k_i+k_R^-+k_R^+ - \sqrt{(k_i+k_R^-+k_R^+)^2 - 4k_i k_R^-}\right).
\end{split}
\end{align}
Finally we are in sight of the original goal. We can generate the
steady-state probability distribution of interest by
differentiating the generating function,
\begin{align}
p(m) = m! \left.\frac{\partial^m}{\partial z^m} f(z) \right|_{z=0},
\end{align}
which follows easily from its definition. Some contemplation reveals
that repeated application of the derivative rule used above will produce
products of the form $\alpha(\alpha+1)(\alpha+2)\cdots(\alpha+m-1)$ in
the expression for $p(m)$ and similarly for $\beta$ and $\gamma$. These
resemble ratios of factorials, but since $\alpha$, $\beta$, and $\gamma$
are not necessarily integer, we should express the ratios using gamma
functions instead. More precisely, one finds
\begin{align}
p(m) = \frac{
        \Gamma(\alpha + m)\Gamma(\beta + m)\Gamma(k_R^+ + k_R^-)
        }
        {
        \Gamma(\alpha)\Gamma(\beta)\Gamma(k_R^+ + k_R^- + m)
        }
\frac{b^m}{m!}{_2F_1}(\alpha+m, \beta+m, k_R^++k_R^-+m; -b)
\label{eq:p_m_bursty+rep}
\end{align}
which is finally the probability distribution we sought to derive.

\subsection{Moments}
We can verify the correctness of~\eref{eq:p_m_bursty+rep}
by comparison with Gillespie simulation.
We can also easily compute moments of the distribution from
the generating function, and check these against raw moments computed
directly from the master equation.

The $j$-th raw moment of the distribution can be computed from the
generating function from the formula
\begin{align}
\langle m^j \rangle
= \left. \left(z \frac{\partial}{\partial z}\right)^j f(z)\right|_{z=1}.
\end{align}
Note that \textit{probabilities} are generated by differentiating
and evaluating at $z=0$, while \textit{moments} are generated by
differentiating and evaluating at $z=1$. This also aligns with the
normalization constraint, which can be thought of as the zeroth raw moment.

The mean mRNA is given by
\begin{align}
\langle m \rangle = \left.z \pderiv[f]{z} \right|_{z=1}
= \frac{\alpha\beta b}{k_R^+ + k_R^-}
= k_i b\frac{k_R^-}{k_R^+ + k_R^-},
\end{align}
which is intuitive and appealingly simple:
the mean mRNA is just the mean mRNA in the absense of repression
($k_i b$, the burst rate time mean burst size)
times the fraction of time the repressor is bound, $k_R^-/(k_R^+ + k_R^-)$.

To compute the variance, let us first compute
\begin{align}
\begin{split}
\langle m^2\rangle
&= \left .z\pderiv{z}\left(z\pderiv[f]{z}\right) \right|_{z=1}
= \left. z\left(\pderiv[f]{z} + z\psecderiv[f]{z}\right)\right|_{z=1}
\\
&= \frac{\alpha\beta b}{k_R^+ + k_R^-}
        + \frac{\alpha(\alpha+1)\beta(\beta+1) b^2}
                {(k_R^+ + k_R^-)(k_R^+ + k_R^- + 1)}
\\
&= \frac{k_i b k_R^-}{k_R^+ + k_R^-}
        + \frac{k_i k_R^- b^2 (1 + k_i + k_R^+ + k_R^- + k_i k_R^-)}
                {(k_R^+ + k_R^-)(k_R^+ + k_R^- + 1)},
\end{split}
\end{align}
from which the variance is
\begin{align}
\begin{split}
var(m) = \langle m^2\rangle - \langle m\rangle^2
&= \frac{k_i b k_R^-}{k_R^+ + k_R^-}
\left(
1 + \frac{b(1 + k_i + k_R^+ + k_R^- + k_i k_R^-)} {k_R^+ + k_R^- + 1}
- \frac{k_i b k_R^-}{k_R^+ + k_R^-}
\right)
\\
&= k_i b\frac{k_R^-}{k_R^+ + k_R^-}
\left(
1 + b + k_i b\frac{1 + k_R^-} {k_R^+ + k_R^- + 1}
- \frac{k_i b k_R^-}{k_R^+ + k_R^-}
\right)
\\
&= k_i b\frac{k_R^-}{k_R^+ + k_R^-}
\left(
1 + b + k_i b\frac{k_R^+}{(k_R^+ + k_R^-)(k_R^+ + k_R^- + 1)}
\right).
\end{split}
\end{align}
The Fano factor is the quantity in parentheses. This can be interpreted as
the variability arising from the negative binomial distribution of the
unregulated promoter $(1+b)$ plus additional noise arising from repression.

\mmnote{Still to writeup: moment derivation direct from CME to compare}

\subsection{Numerical considerations and recursion formulas}
We would like to carry out Bayesian parameter inference on FISH data
from~\cite{Jones2014}, using~\eref{eq:p_m_bursty+rep} as our
likelihood. This requires accurate (and preferably fast)
numerical evaluation of the hypergeometric function ${_2F_1}$,
which is a notoriously hard problem~\cite{Pearson2016, Gil2007},
and our particular needs here present an especial challenge as we show below.

The hypergeometric function is defined by its Taylor series as
\begin{align}
{_2F_1}(a,b,c;z) 
= \sum_{l=0}^\infty
\frac{\Gamma(a + l)\Gamma(b + l)\Gamma(c)}
        {\Gamma(a)\Gamma(b)\Gamma(c + l)}
\frac{z^l}{l!}
\end{align}
for $|z|<1$, and by analytic continuation elsewhere.
If $z\lesssim1/2$ and $\alpha$ and $\beta$ are not too large
(absolute value below 20 or 30),
then the series converges quickly and an accurate numerical representation is
easily computed by truncating the series after a reasonable number of terms.
Unfortunately, we need to evaluate ${_2F_1}$ over mRNA copy numbers fully out
to the tail of the distribution, which can easily reach 50, possibly 100.
From~\eref{eq:p_m_bursty+rep}, this means evaluating ${_2F_1}$
repeatedly for values of $a$, $b$, and $c$ spanning the full range
from $\mathcal{O}(1)$ to $\mathcal{O}(10^2)$,
even if $\alpha$, $\beta$, and $\gamma$
in~\eref{eq:p_m_bursty+rep} are small,
with the situation even worse if they are not small.
A naive numerical evaluation of the series definition will be
prone to overflow and, if any of $a,b,c<0$, then some successive terms in the
series have alternating signs which can lead to catastrophic cancellations.

The solution involves recursion relations originally explored by Gauss,
and studied extensively in~\cite{Pearson2016, Gil2007}.

%%%%%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix

% \bibliographystyle{nature}
\bibliographystyle{abbrv}
\bibliography{../library}

\end{document}