### 2020-05-19 doc/abstract.tex
initial splitting of document into several files

% !TEX root = ./busty_transcription.tex
\begin{abstract}
Abstract here.
\end{abstract}