### 2020-05-19 doc/appendix_01_bayesian_inference.tex

% !TEX root = ./busty_transcription.tex
\section{Bayesian inference}
\mmnote{
\begin{itemize}
\item Bayes theorem.
Elaborate on posterior as plausibility over
space of model parameters.
\item Choosing likelihoods.
Easy here, just the steady-state distributions of our master
equations, assuming iid single-cell measurements.
\item Choosing priors.
Not critical here because we are data-rich and nonhierarchical,
as long as we don't exclude ground truth. In other applications,
prior can matter a lot. This is a feature, not a bug: it's
revealing hidden structure that was implied by your modeling
assumptions and you just didn't realize it. Predictive checks to
verify you're not excluding any plausible values.
\item Marginalization/expectations.
Analytically hard, trivial with sampling.
\item MCMC.
Lot's of fun here, and deep analogies to stat mech and
Hamiltonian mechanics worth exploring. In a perfect world, you'd
wave a magic wand and generate independent draws from target
posterior. But we don't know how. Next best: generate a Markov
chain of $N$ \textit{correlated} samples, estimate an
autocorrelation time $\tau$, then you've got roughly $N/\tau$
effective samples. Obviously with some assumptions.
\item Posterior predictive checks.
So the posterior is fine, but is the model actually not wrong?
PPC is essential to verify that your model is a plausible
generative process for the data.
\item Discuss identifiability, degeneracy, and failed PPCs in main text. Better thru example anyway.
\end{itemize}
}