### 2020-05-19 doc/section_02_means.tex

% !TEX root = ./busty_transcription.tex
\section{Means}
\subsection{Fold-changes are indistinguishable across models}
As a first stop on our search for the ``right'' model of simple repression, let
us consider what we can learn from theory and measurements of fold-change in
mean expression. The fold-change is defined as
\begin{equation}
\text{fold-change}
= \frac{\langle \text{gene expression with repressor present} \rangle}
        {\langle \text{gene expression with repressor absent} \rangle}
= \frac{\langle m\rangle(R\ne 0)}{\langle m\rangle(R = 0)}
\label{eq:fc_def}
\end{equation}
where angle brackets denote the averge over a population of cells and mean mRNA
$\langle m\rangle$ is viewed as a function of repressor copy number $R$. The
second equality in~\ref{eq:fc_def} follows from assuming that the translation
efficiency, i.e., the number of proteins translated per mRNA, is the same in
both conditions. In other words, we assume that mean protein level is
proportional to mean mRNA level, and that the proportionality constant is the
same in both conditions and therefore cancels out in the ratio. This is
reasonable since the cells in the two conditions are identical except for the
presence of the transcription factor, and it is difficult to imagine the
transcription factor having any direct effect on translation.

Fold-change has proven a very convenient observable in past
work~\cite{Garcia2011a, Razo-Mejia2018, Chure2019}. Part of its utility in
dissecting transcriptional regulation is its ratiometric nature, which is always
easier than making an absolute measurement. Also, by measuring otherwise
identical cells with and without a transcription factor present, any biological
noise common to both conditions can be made to cancel away. In addition, the
presence of different operator sites is known to change the basal expression
level in the absence of repressor~\mmnote{look up cite}. Computing fold-changes
neatly cancels this effect as well.

\fig{fig1:means_cartoons} depicts a smorgasbord of mathematized cartoons for
simple repression that have appeared in previous literature. For each cartoon,
we calculate the fold-change in mean gene expression as predicted by that model,
deferring some algebraic details to the appendix. What we will find is that all
cartoons collapse to a single master curve, shown in \fig{fig1:means_cartoons}c,
which contains just two parameters. We label the parameters $\Delta F_R$, an
effective free energy parametrizing the repressor-DNA interaction, and $\rho$,
which subsumes all details of transcription in the absence of repressors. We
will offer some intuition for why this master curve exists and discuss why it is
likely impossible to discriminate ``right'' from ``wrong'' cartoons given only
measurements of fold-changes in expression.

\subsubsection{Equilibrium, 2-state}
In this simplest model, the promoter is idealized as existing in one of two
states, either repressor bound or repressor unbound. The rate of transcription,
and therefore the level of gene expression, is assumed to be proportional to the
fraction of time spent in the repressor unbound state. Mathematically, this
amounts to the statement that the mean mRNA level is given by
\begin{equation}
\langle m \rangle = \frac{r}{\gamma}
        \left(1 + \frac{R}{N_{NS}} e^{-\beta\Delta\epsilon_R}\right)^{-1},
\end{equation}
where $r$ is the transcription rate from the repressor unbound state, $\gamma$
is the mRNA degradation rate, $R$ is repressor copy number, $N_{NS}$ is the
number of nonspecific binding sites in the genome (where repressors spend most
of their time when not bound to the operator, their specifc binding site),
$\beta=1/k_BT$, and $\Delta\epsilon_R$ is the binding energy of a repressor to
its operator site. The derivation of this result is deferred to the appendix.

The fold-change is easily found as the ratio of mean mRNA with and without
repressor. This gives
\begin{equation}
\text{fold-change}
= \left(1 + \frac{R}{N_{NS}} e^{-\beta\Delta\epsilon_R}\right)^{-1},
\end{equation}
which clearly matches the form of the master curve in~\eq{fig1:means_cartoons}
with $\rho=1$.

In fact it was noted in~\cite{Chure2019} that this two-state model can be viewed
as the coarse-graining of any equilibrium promoter model in which no
transcriptionally active states have transcription factor bound. Put
differently, there is no overlap between TF bound states and transcriptionally
active states. We will see this explicitly in the 3-state equilibrium model
below, but perhaps surprising is that something analogous carries over even to
the nonequilibrium models we consider later.

\subsubsection{Equilibrium, 3-state}
Compared to the previous model, here we fine-grain the repressor unbound state
into two separate states: empty, and RNAP bound. This picture was used
in~\cite{Garcia2011a} as we use it here, and in~\cite{Razo-Mejia2018}
and~\cite{Chure2019} it was generalized to incorporate small-molecule inducers
that bind the repressor. Interestingly, the effect of this is simply to
renormalize $R$ from the total number of repressors to a smaller effective
number, roughly (though not exactly) the number which are unbound by inducers.

The mean mRNA, as derived in the appendix from a similar enumeration of states
and weights as the previous model, is
\begin{equation}
\langle m \rangle = \frac{r}{\gamma}
\frac{\frac{P}{N_{NS}} e^{-\beta\Delta\epsilon_P}}
        {
        1 + \frac{R}{N_{NS}} e^{-\beta\Delta\epsilon_R}
        + \frac{P}{N_{NS}} e^{-\beta\Delta\epsilon_P}
        },
\end{equation}
where the new variables are $\Delta\epsilon_P$, the difference in RNAP binding
energy to its specific site (the promoter) relative to an average nonspecific
background site, and the RNAP copy number, $P$. The fold-change again follows
immediately as
\begin{align}
\text{fold-change}
&= \frac{\frac{P}{N_{NS}} e^{-\beta\Delta\epsilon_P}}
        {
        1 + \frac{R}{N_{NS}} e^{-\beta\Delta\epsilon_R}
        + \frac{P}{N_{NS}} e^{-\beta\Delta\epsilon_P}
        }
\frac{1 + \frac{P}{N_{NS}} e^{-\beta\Delta\epsilon_P}}
        {\frac{P}{N_{NS}} e^{-\beta\Delta\epsilon_P}}
\\
&= \left(
1 + \frac{\frac{R}{N_{NS}} e^{-\beta\Delta\epsilon_R}}
        {1 + \frac{P}{N_{NS}} e^{-\beta\Delta\epsilon_P}}
\right)^{-1}
\\
&= (1 + \exp(-\Delta F_R - \log\rho))^{-1},
\end{align}
with $\Delta F_R = \beta\Delta\epsilon_R - \log(R/N_{NS})$ and $\rho = 1 +
\frac{P}{N_{NS}}\mathrm{e}^{-\beta\Delta\varepsilon_P}$ as shown
in~\fig{fig1:means_cartoons}.

\begin{figure}%[h!]
\centering
\includegraphics[width=\textwidth]{../figures/fig1/fig1.ai}
\caption{\textbf{An overview of the simple repression motif at the level of
means.} (A) depicts the qualitative biological picture of simple repression. (B)
shows a variety of possible mathematized cartoons of simple repression, along
with the effective parameter $\rho$ which subsumes all regulatory details of the
architecture excluding the repressor. (C) shows the ``master curve'' to which
all cartoons in (B) collapse.}
\label{fig1:means_cartoons}
\end{figure}

\subsubsection{Nonequilibrium model 1 - Poisson promoter}
For our first kinetic model, we consider the simplest possible picture with only
two states, repressor bound and unbound. This is exactly the model used for the
main results of~\cite{Jones2014}. In this picture, repressor association and
dissociation rates from its operator site, $k_R^+$ and $k_R^-$, respectively,
govern transitions between the two states. When the system is in the unbound
state, transcriptions initiate at rate $r$, which represents a coarse-graining
of all the downstream processes into a single effective rate. mRNA is again
degraded at rate $\gamma$.

Let $p_{m,R}(t)$ denote the joint probability of finding the system in the
repressor bound state with $m$ mRNA molecules present at time $t$, and similarly
define $p_{m,U}(t)$ for the repressor unbound state. This model is governed by a
master equation giving the time evolution of $p_{m,R}(t)$ and $p_{m,U}(t)$,
which is
\begin{align}
\begin{split}
\deriv{t}p_{m,R}(t) =& - k_R^- p_{m,R}(t) + k_R^+ p_{m,U}(t)
                + (m+1)\gamma p_{m+1,R}(t) - \gamma p_{m,R}(t)
\\
\deriv{t}p_{m,U}(t) =&\; k_R^- p_{m,R}(t) - k_R^+ p_{m,U}(t)
                        + rp_{m-1,U}(t) - rp_{m,U}(t)
                \\
                &+ (m+1)\gamma p_{m+1,U}(t) - \gamma p_{m,U}(t).
\label{eq:poisson_promoter_cme}
\end{split}
\end{align}
In each equation, the first two terms describe transitions between promoter
states due to repressors unbinding and binding, respectively. The final two
terms describe degradation of mRNA, decreasing the copy number by one, and the
terms with coefficient $r$ describe transcription initiation increasing the mRNA
copy number by one.

We can greatly simplify the notation, which will be especially useful for the
more complicated models yet to come, by re-expressing the master equation in
vector form. The promoter states are collected into a vector and the rate
constants are collected into matrices as
\begin{equation}
\vect{p}_m = \begin{pmatrix} p_{m,R} \\ p_{m,U} \end{pmatrix},\
\matr{K} = \begin{pmatrix} -k_R^- & k_R^+ \\ k_R^- & -k_R^+ \end{pmatrix},\
\matr{R} = \begin{pmatrix} 0 & 0 \\ 0 & r \end{pmatrix},\
\label{eq:2state_cme_matrices}
\end{equation}
so that the master equation may be condensed as
\begin{equation}
\deriv{t}\vect{p}_m(t) =
\left( \matr{K} - \matr{R} - \gamma m \matr{I} \right) \vect{p}_m(t)
                + \matr{R} \vect{p}_{m-1}(t)
                + \gamma (m+1) \matr{I} \vect{p}_{m+1}(t),
\label{eq:2state_rep_cme}
\end{equation}
where $\matr{I}$ is the identity matrix. Taking steady state by setting time
derivatives to zero, the mean mRNA can be found to be
\begin{equation}
\langle m \rangle = \frac{r}{\gamma}
        \left(1 + \frac{k_R^+}{k_R^-}\right)^{-1},
\end{equation}
with the algebra details again deferred to the appendix. Recall $k_R^+$ is
proportional to the repressor copy number, so in computing fold-change, absence
of repressor corresponds to $k_R^+\rightarrow0$. Therefore fold-change in this
model is simply
\begin{equation}
\text{fold-change} = \left(1 + \frac{k_R^+}{k_R^-}\right)^{-1},
\end{equation}
again matching the master curve of~\fig{fig1:means_cartoons} with $\rho=1$.

\subsubsection{Nonequilibrium model 2 - RNAP bound/unbound states}
Analogous to the three-state equilibrium model, this model fine-grains the
repressor unbound state of nonequilibrium model 1, resolving it into an empty
promoter state and an RNAP-bound state. Note in this picture, in contrast with
model 4 below, transcription initiation is accompanied by a promoter state
change, in keeping with the interpretation as RNAP-bound and empty states: if an
RNAP successfully escapes the promoter and proceeds to elongation of a
transcript, clearly it is no longer bound at the promoter! Therefore another
RNAP must bind before another transcript can be initiated.

The master equation governing this model is analogous
to~\eqrange{eq:2state_cme_matrices}{eq:2state_rep_cme} for model 1 above. The
main subtlety arises since transcription initiation accompanies a promoter state
change. This can be understood by analogy to $\matr{K}$. The off-diagonals and
diagonals of $\matr{K}$ correspond to transitions arriving at or departing from,
respectively, the promoter state of interest. If transcription initiations are
accompanied by promoter state changes, we must have separate matrices for
arriving and departing transcription events since the arriving and departing
transitions have different initial copy number of mRNA, unlike for $\matr{K}$
where they are the same. The master equation for this model is
\begin{equation}
\deriv{t}\vect{p}_m(t) =
\left( \matr{K} - \matr{R_D} - \gamma m \matr{I} \right) \vect{p}_m(t)
                + \matr{R_A} \vect{p}_{m-1}(t) +
                \gamma (m+1) \matr{I} \vect{p}_{m+1}(t),
\label{eq:3state_rep_cme}
\end{equation}
with the state vector and promoter transition matrix defined as
\begin{equation}
\vect{p}_m = \begin{pmatrix} p_{m,R} \\ p_{m,E} \\ p_{m,P} \end{pmatrix},\
\matr{K} = \begin{pmatrix} -k_R^- & k_R^+ & 0 \\
                        k_R^- & -k_R^+ -k_P^+ & k_P^- \\
                        0 & k_P^+ & -k_P^- 
                \end{pmatrix},
\label{eq:3state_cme_matrices_pt1}
\end{equation}
and the initiation matrices given by
\begin{equation}
\matr{R_A} = \begin{pmatrix}
                0 & 0 & 0 \\ 0 & 0 & r \\ 0 & 0 & 0
                \end{pmatrix},\
\matr{R_D} = \begin{pmatrix}
                0 & 0 & 0 \\ 0 & 0 & 0 \\ 0 & 0 & r
                \end{pmatrix},
\label{eq:3state_cme_matrices_pt2}
\end{equation}
The elements of $\vect{p}_m$ encode the probabilities of having $m$ mRNA present
along with the promoter having repressor bound, being empty, or having RNAP
bound, respectively. $\matr{R_A}$ describes probability flux \textit{arriving}
at the state $\vect{p}_m$ from a state with one fewer mRNA, namely
$\vect{p}_{m-1}$, and $\matr{R_D}$ describes probability flux \textit{departing}
from the state $\vect{p}_m$ for a state with one more mRNA, namely
$\vect{p}_{m+1}$. $\matr{K}$ is closely analogous to model 1.

Mean mRNA at steady state is found analogously to model 1, with the result
\begin{equation}
\langle m\rangle = \frac{r}{\gamma}
        \frac{k_R^- k_P^+}
        {k_R^- k_P^+ + k_R^- (k_P^- + r) + k_R^+ (k_P^- + r)}.
\end{equation}
Fold-change is again found from the ratio prescribed by~\eq{eq:fc_def}, from
which we have
\begin{align}
\text{fold-change}
&=      \frac{k_R^- k_P^+}
        {k_R^- k_P^+ + k_R^- (k_P^- + r) + k_R^+ (k_P^- + r)}
        \frac{k_P^+ + k_P^- + r}{k_P^+}
\\
&=      \left(1 + \frac{k_R^+}{k_R^-}
                \frac{k_P^- + r}{k_P^+ + k_P^- + r}
        \right)^{-1}
\\
&=      \left(1 + \frac{k_R^+}{k_R^-}
        \left(1 + \frac{k_P^+}{k_P^- + r}\right)^{-1}
        \right)^{-1},
\end{align}
which follows the master curve with $\rho = 1 + k_P^+/(k_P^- + r)$ as claimed.

\subsubsection{Nonequilibrium model 3 - Multistep transcription initiation and
escape} One might reasonably complain that the first two ``nonequilibrium''
models we have considered are straw men. Their steady states necessarily satisfy
detailed balance which is equivalent to thermodynamic equilibrium. Why is this
the case? At steady state there is by definition no net probability flux in or
out of each promoter state, but since the promoter states form a linear chain,
there is only one way in or out of the repressor bound and RNAP bound states,
implying each edge must actually have a net zero probability flux, which is the
definition of detailed balance (usually phrased as equality of forward and
reverse transition fluxes).

Now we consider model 3 in~\eq{fig1:means_cartoons} which allows the possibility
of true nonequilibrium steady-state fluxes through the promoter states. We point
out that this model was considered previously in~\cite{Mitarai2015} where a
comparison was made with model 1 as used in~\cite{Jones2014}. The authors
of~\cite{Mitarai2015} argued that the additional complexity is essential to
properly account for the noise in the mRNA distribution. We will weigh in on
both models later when we consider observables beyond fold-change.

The master equation governing this model is identical in form to model 2 above,
namely
\begin{equation}
\deriv{t}\vect{p}_m(t) =
\left( \matr{K} - \matr{R_D} - \gamma m \matr{I} \right) \vect{p}_m(t)
                + \matr{R_A} \vect{p}_{m-1}(t) +
                \gamma (m+1) \matr{I} \vect{p}_{m+1}(t),
\end{equation}
but obviously with a higher-dimensional state space and different matrices. The
state vector and promoter transition matrix are now
\begin{equation}
\vect{p}_m = \begin{pmatrix} p_{m,R} \\ p_{m,E} \\
                             p_{m,C} \\ p_{m,O}\end{pmatrix},\
\matr{K} = \begin{pmatrix} -k_R^- & k_R^+ & 0 & 0\\
                        k_R^- & -k_R^+ -k_P^+ & k_P^- & 0 \\
                        0 & k_P^+ & -k_P^- - k_O & 0 \\
                        0 & 0 & k_O & 0
                \end{pmatrix},
\end{equation}
with the four promoter states, in order, being repressor bound, empty, RNAP
closed complex, and RNAP open complex. Besides increasing dimension by one
again, the only new feature in $\matr{K}$ is the rate $k_O$, representing the
(irreversible) rate of open complex formation from the closed complex. The
initiation matrices are given by
\begin{equation}
\matr{R_A} = \begin{pmatrix}
        0 & 0 & 0 & 0 \\ 0 & 0 & 0 & r \\ 0 & 0 & 0 & 0\\ 0 & 0 & 0 & 0
                \end{pmatrix},\
\matr{R_D} = \begin{pmatrix}
        0 & 0 & 0 & 0 \\ 0 & 0 & 0 & 0 \\ 0 & 0 & 0 & 0\\ 0 & 0 & 0 & r
                \end{pmatrix},
\end{equation}
again closely analogous to model 2.

The expression for mean mRNA is substantially more complicated now, as worked
out in the appendix where we find
\begin{equation}
\langle m\rangle = \frac{r}{\gamma}
        \frac{k_R^- k_P^+ k_O}
        {k_R^- [(k_P^+ (k_O + r) + r(k_P^- + k_O)] + k_R^+ r(k_P^- + k_O)},
\end{equation}
which can be simplified to
\begin{equation}
\langle m\rangle
= \frac{r}{\gamma}
\frac{\frac{k_P^+ k_O}{r(k_O + k_P^-)}}
        {1 + \frac{k_P^+ (k_O + r)}{r(k_O + k_P^-)} + \frac{k_R^+}{k_R^-}}.
\end{equation}
The strategy is to isolate the terms involving the repressor, so that now the
fold-change is seen to be simply
\begin{align}
\text{fold-change}
&= \frac{\frac{k_P^+ k_O}{r(k_O + k_P^-)}}
        {1 + \frac{k_P^+ (k_O + r)}{r(k_O + k_P^-)} + \frac{k_R^+}{k_R^-}}
        \frac{1 + \frac{k_P^+ (k_O + r)}{r(k_O + k_P^-)}}
                {\frac{k_P^+ k_O}{r(k_O + k_P^-)}}
\\
&= \left(
        1 + \frac{k_R^+}{k_R^-}
        \left(1 + \frac{k_P^+ (k_O + r)}{r(k_O + k_P^-)}\right)^{-1}
\right)^{-1},
\end{align}
surprisingly reducing to the master curve once again, with $\rho = 1 +
\frac{k_P^+ (k_O + r)}{r(k_O + k_P^-)}$.

This example hints that an arbitrarily fine-grained model of downstream
transcription steps may still be collapsed to the form of the master curve
in~\fig{fig1:means_cartoons}B so long as the repressor binding is exclusive with
transcriptionally active states. We offer this as a conjecture, and we suspect
that a careful argument using the King-Altman diagram method~\cite{King1956,
Hill1966} might furnish a ``proof.'' Our focus here is not on full generality
but rather to survey an assortment of plausible models that have been proposed
in the literature for simple repression.

\subsubsection{Nonequilibrium model 4 - ``Active'' and ``inactive'' states}
This model is at the core of the theory in~\cite{Razo-Mejia2020}. At a glance
the cartoon for this model may appear very similar to model 2, and
mathematically it is, but the interpretation is rather different. In model 2, we
interpreted the third state literally as an RNAP-bound promoter and modeled
initiation of a transcript as triggering a promoter state change, making the
hopefully uncontroversial assumption that an RNAP can only make one transcript
at a time. In contrast, in the present model the promoter state does
\textit{not} change when a transcript is initiated. So we no longer interpret
these states as literally RNAP bound and unbound but instead as coarse-grained
``active'' and ``inactive'' states, the details of which we leave unspecified
for now. We will comment more on this model below when we discuss Fano factors
of models.

Mathematically this model is very similar to models 1 and 2. Like model 1, the
matrix $R$ describing transcription initiation is diagonal, namely
\begin{equation}
\matr{R} = \begin{pmatrix}
                0 & 0 & 0 \\ 0 & 0 & 0 \\ 0 & 0 & r
        \end{pmatrix}.
\end{equation}
so the master equation takes verbatim the same form as it did for model
1,~\eq{eq:2state_rep_cme}. Meanwhile the promoter transition matrix $K$ is the
same as~\eq{eq:3state_cme_matrices_pt1} from model 2, although we relabel the
rate constants from $k_P^\pm$ to $k^\pm$ to reiterate that these are not simply
RNAP binding and unbinding rates.

Carrying out the algebra, the mean mRNA can be found to be
\begin{equation}
\langle m\rangle = \frac{r}{\gamma}
\frac{k_R^- k^+}
{k_R^- k^+ + k_R^- k^- + k_R^+ k^-},
\end{equation}
and the fold-change readily follows,
\begin{align}
\text{fold-change}
&=      \frac{k_R^- k^+}{k_R^- k^+ + k_R^- k^- + k_R^+ k^-}
        \frac{k_R^- k^+ + k_R^- k^-}{k_R^- k^+}
\\
&=      \left(1 + \frac{k_R^+}{k_R^-}
                \left(1 + \frac{k^+}{k^-}\right)^{-1}
        \right)^{-1},
\end{align}
from which we see $\rho = 1 + k^+/k^-$ as shown in~\fig{fig1:means_cartoons}.

\subsubsection{Nonequilibrium model 5 - Bursty promoter}
We will have much more to say about this model in subsequent sections. For now
we sketch it as an intuitive analog to model 1, with just two states, repressor
bound or unbound, and transition rates between them of $k_R^+$ and $k_R^-$. In
model 1, when in the unbound state, single mRNA transcripts are produced as a
Poisson process with some characteristic rate $r$. This model by contrast
produces, at some Poisson rate $k_i$, \textit{bursts} of transcripts with a
geometrically distributed burst size and a mean burst size $b$.

From this intuitive picture and by analogy to model 1, then, it should be
plausible that the mean mRNA level is
\begin{equation}
\langle m\rangle = \frac{k_i b}{\gamma}
        \left(1 + \frac{k_R^+}{k_R^-}\right)^{-1},
\end{equation}
which will turn out to be correct from a careful calculation. For now, we simply
note that just like model 1, the fold-change becomes
\begin{equation}
\text{fold-change} = \left(1 + \frac{k_R^+}{k_R^-}\right)^{-1}
\end{equation}
with $\rho=1$ also like model 1.

We will also see later how this model emerges as a natural limit of model 4.

\subsubsection{Summary and discussion of results across models for fold-changes in mean expression}
What have we learned from our exploration of the models
in~\fig{fig1:means_cartoons} and their predictions of fold-changes in mean
expression levels? The takeaway is the existence of a master curve
(\fig{fig1:means_cartoons}C) to which the fold-change predictions of all the
models collapse. This master curve is parametrized by only two effective
parameters: $\Delta F_R$, which characterizes the repressor, and $\rho$, which
characterizes all other features of the promoter architecture. The key
assumption underpinning this result is that zero transcription occurs when a
repressor is bound to its operator. Note, however, that we are agnostic about
the molecular mechanism which achieves this; steric effects are the obvious
mechanism, but, for instance, ``action-at-a-distance'' mediated by kinked DNA
due to repressors bound tens or hundreds of nucleotides upstream of a promoter
is plausible as well.

Why does the master curve of~\fig{fig1:means_cartoons}C exist at all? This
brings to mind the deep questions posed in, e.g.,~\cite{Frank2013}
and~\cite{Frank2014a}, asking us to search for general ``symmetries'' of the
problem instead of specifying generative models of the known microscopics of a
system. In our case, the key feature seems to be the exclusive nature of
repressor and RNAP binding, which allows the parameter describing the repressor,
$\Delta F_R$, to cleanly separate from all other details of the promoter
architecture, which are encapsulated in $\rho$. In the graph-theoretic language
which was applied to similar problems in~\cite{Gunawardena2012}
and~\cite{Ahsendorf2014}, the repressor-bound state is a leaf of the graph
describing transitions between promoter states. Arbitrary nonequilibrium
behavior can occur on the rest of the promoter state space, but it may all be
swept up in the effective parameter $\rho$, to which the repressor makes no
contribution.

As suggested in~\cite{Chure2019}, we believe this master curve should generalize
to architectures with multiple repressor binding sites, as long as the
exclusivity of transcription factor binding and transcription initation is
maintained. The interpretation of $\Delta F_R$ is then of an effective free
energy of all repressor bound states. In an equilibrium picture this is simply
given by the log of the sum of Boltzmann weights of all repressor bound states,
which looks like the log of a partition function of a subsystem. In a
nonequilibrium picture, while we can still mathematically gather terms and give
the resulting collection the label $\Delta F_R$, it is unclear if the physical
interpretation as an effective free energy makes sense. As is well known, free
energies cannot be assigned unambiguously to states out of equilibrium because
the free energy change along a generic path traversing the state space is path
dependent, unlike at equilibrium. A consequence of this is that, out of
equilibrium, $\Delta F_R$ is no longer a simple sum of Boltzmann weights.
Instead it resembles a restricted sum of King-Altman diagrams~\cite{King1956,
Hill1966}. Following the work of Hill~\cite{Hill1989}, it may yet be possible to
interpret this expression as an effective free energy, but this remains unclear
to us. We leave this an open problem for future work.

If we relax the requirement of exclusive repressor-RNAP binding, one could
imagine models in which repressor and RNAP doubly-bound states are allowed,
where the repressor's effect is to \text{reduce} the transcription rate rather
than setting it to zero. Our results do not strictly apply to such a model,
although we note that if the repressor's reduction of the transcription rate is
substantial, such a model might still be well-approximated by one of the models
in~\fig{fig1:means_cartoons}.

One may worry that our ``one curve to rule them all'' is a mathematical
tautology. In fact we \textit{agree} with this criticism if $\Delta F_R$ is
``just a fitting parameter'' and cannot be meaningfully interpreted as a real,
physical free energy. An analogy to Hill functions is appropriate here. One of
their great strengths and weaknesses, depending on the use they are put to, is
that their parameters coarse-grain many details and are generally not
interpretable in terms of microscopic models, for deep reasons discussed at
length in~\cite{Frank2013}. By contrast, our master curve claims to have the
best of both worlds: a coarse-graining of all details besides the repressor into
a single effective parameter $\rho$, while simultaneously retaining an
interpretation of $\Delta F_R$ as a physically meaningful and interpretable free
energy. Our task, then, is to prove or disprove this claim.

How do we test this and probe the theory with fold-change measurements? There is
a fundamental limitation in that the master curve is essentially a one-parameter
function of $\Delta F_R + \log\rho$. Worse, there are many \textit{a priori}
plausible microscopic mechanisms that could contribute to the value of $\rho$,
such as RNAP binding and escape kinetics~\cite{DeHaseth1998, Mitarai2015},
and/or supercoiling accumulation and release~\cite{Chong2014, Sevier2016},
and/or, RNAP clusters analogous to those obvserved in
eukaryotes~\cite{Cisse2013, Cho2016} and recently also observed in
bacteria~\cite{Ladouceur2020}. Even if $\Delta F_R$ is measured to high
precision, inferring the potential microscopic contributions to $\rho$, buried
inside a log no less, from fold-change measurements seems hopeless. As a
statistical inference problem it is entirely nonidentifiable, in the language
of~\cite{Gelman2013}, Section 4.3.

If we cannot simply infer values of $\rho$ from measurements of fold-change, can
we perturb some of the parameters that make up $\rho$ and measure the change?
Unfortunately we suspect this is off-limits experimentally: most of the
potential contributors to $\rho$ are global processes that affect many or all
genes. For instance, changing RNAP association rates by changing RNAP copy
numbers, or changing supercoiling kinetics by changing topoisomerase copy
numbers, would massively perturb the entire cell's physiology and confound any
determination of $\rho$.

One might instead imagine a bottom-up modeling approach, where we mathematize a
model of what we hypothesize the important steps are and are not, use in vitro
data for the steps deemed important, and \textit{predict} what $\rho$ should be.
But again, because of the one-parameter nature of the master curve, many
different models will likely make indistinguishable predictions, and without any
way to experimentally perturb in vivo, there is no clear way to test whether the
modeling assumptions are correct.

In light of this, we prefer the view that parameters and rates are not directly
comparable between cartoons in~\fig{fig1:means_cartoons}. Rather, parameters in
the simpler cartoons represent coarse-grained combinations of parameters in the
finer-grained models. For instance, by equating $\rho$ between any two models,
one can derive various possible correspondences between the two models'
parameters. Note that these correspondences are clearly not unique, since many
possible associations could be made. It then is a choice as to what microscopic
interpretations the model-builder prefers for the parameters in a particular
cartoon, and as to which coarse-grainings lend intuition and which seem
nonsensical. Indeed, since it remains an open question what microscopic features
dominate $\rho$ (as suggested above, perhaps RNAP binding and escape
kinetics~\cite{DeHaseth1998, Mitarai2015}, or supercoiling accumulation and
release~\cite{Chong2014, Sevier2016}, or, something more exotic like RNAP
clusters~\cite{Cisse2013, Cho2016, Ladouceur2020}), we are hesitant to put too
much weight on any one microscopic interpretation of model parameters that make
up $\rho$.

One possible tuning knob to probe $\rho$ that would not globally perturb the
cell's physiology is to manipulate RNAP binding sites. Work such
as~\cite{Kinney2010} has shown that models of sequence-dependent RNAP can be
inferred from data, and the authors of~\cite{Brewster2012} showed that the model
of~\cite{Kinney2010} has predictive power by using the model to \textit{design}
binding sites of a desired affinity. But for our purposes, this begs the
question: the authors of~\cite{Kinney2010} \textit{assumed} a particular model
(essentially our 3-state equilibrium model but without the repressor), so it is
unclear how or if such methods can be turned around to \textit{compare}
different models of promoter function.

We have already pointed out that the master curve of~\fig{fig1:means_cartoons}
is essentially a one-parameter model, the one parameter being $\Delta F_R +
\log\rho$. By now the reader may be alarmed as to how can we even determine
$\Delta F_R$ and $\rho$ independently of each other, never mind shedding a lens
on the internal structure of $\rho$ itself. A hint is provided by the weak
promoter approximation, invoked repeatedly in prior studies~\cite{Bintu2005c,
Garcia2011a, Razo-Mejia2018} of simple repression using the 3-state equilibrium
model in~\fig{fig1:means_cartoons}. In that picture, the weak promoter
approximation means $\frac{P}{N_{NS}}\exp(-\beta\Delta\epsilon_P) \ll 1$,
meaning therefore $\rho\approx1$. This is suggestive, but how can we be sure
that $\rho$ is not, for instance, actually $10^2$ and that $\Delta F_R$ hides a
compensatory factor? A resolution is offered by an independent inference of
$\rho$ in the absence of repressors. This was done in~\cite{Razo-Mejia2020} by
fitting nonequilibrium model 4  in~\fig{fig1:means_cartoons}, with zero
repressor (looking ahead, this is equivalent to model 4
in~\fig{fig:constit_cartoons}), to FISH data from~\cite{Brewster2014}. This
provided a determination of $k^+$ and $k^-$, from which their ratio is estimated
to be no more than a few $10^{-1}$ and possibly as small as $10^{-2}$.

The realization that $\rho\approx1$ to an excellent approximation,
\textit{independent} of which model in~\fig{fig1:means_cartoons} one prefers,
goes a long way towards explaining the surprising success of equilibrium models
of simple repression. Even though our 2- and 3-state models get so many details
of transcription wrong, it does not matter because fold-change is a cleverly
designed ratio. Since $\rho$ subsumes all details except the repressor, and
$\log\rho\approx0$, fitting these simple models to fold-change measurements can
still give a surprisingly good estimate of repressor binding energies. So the
ratiometric construction of fold-change fulfills its intended purpose of
canceling out all features of the promoter architecture except the repressor
itself. Nevertheless it is perhaps surprising how effectively it does so:
\textit{a priori}, one might not have expected $\rho$ to be quite so close to 1.

We would also like to highlight the relevance of~\cite{Landman2019} here.
Landman et.\ al.\ reanalyzed and compared in vivo and in vitro data on the lacI
repressor's binding affinity to its various operator sequences. (The in vivo
data was from, essentially, fitting our master curve to expression
measurements.) They find broad agreement between the in vitro and in vivo
values. This reinforces the suspicion that the equilibrium $\Delta\epsilon_R$
repressor binding energies do in fact represent real physical free energies.
Again, \textit{a priori} this did not have to be the case, even knowing that
$\rho\approx1$.

In principle, if $\Delta F_R$ can be measured to sufficient precision, then
deviations from $\rho=1$ become a testable matter of experiment. In practice, it
is probably unrealistic to measure repressor rates $k_R^+$ or $k_R^-$ or
fold-changes in expression levels (and hence $\Delta\epsilon_R$) precisely
enough to detect the expected tiny deviations from $\rho=1$. We can estimate the
requisite precision in $\Delta F_R$ to resolve a given $\Delta\rho$ by noting,
since $\rho\approx1$, that $\log(1+\Delta\rho)\approx \Delta\rho$, so
$\Delta(\Delta F_R) \approx \Delta\rho$. Suppose we are lucky and $\Delta\rho$
happens to be $\sim0.1$, on the high end of our range estimated above. A
determination of $\Delta\epsilon_R/k_BT$ with an uncertainty of barely 0.1 was
achieved in the excellent measurements of~\cite{Razo-Mejia2018}, so this
requires a very difficult determination of $\Delta F_R$ for a very crude
determination of $\rho$, which suggests, to put it lightly, this is not a
promising path to pursue experimentally. It is doubtful that inference of
repressor kinetic rates would be any easier.\footnote{\mmnote{In fact the
uncertainties quoted in Hammar et.\ al.~\cite{Hammar2014} from Elf's group are
small enough to do this calculation, but HJ \& I think their error bars are
insanely optimistic. Not sure I want to blatantly call them out on it though??}}

Moving forward, we have weak evidence supporting the interpretation of $\Delta
F_R$ as a physically real free energy~\cite{Landman2019} and other work casting
doubt~\cite{Hammar2014}. How might we resolve the confusion? If there is no
discriminatory power to test the theory and distinguish the various models with
measurements of fold-changes in means, how do we probe the theory? Clearly to
discriminate between the nonequilibrium models in~\fig{fig1:means_cartoons}, we
need to go beyond means to ask questions about kinetics, noise and even full
distributions of mRNA copy numbers over a population of cells. If the
``one-curve-to-rule-them-all'' is more than a mathematical tautology, then the
free energy of repressor binding inferred from fold-change measurements should
agree with repressor binding and unbinding rates. In other words, the equilbrium
and nonequilibrium definitions of $\Delta F_R$ must agree, meaning
\begin{equation}
\Delta F_R = \beta\Delta\epsilon_R - \log(R/N_{NS})
        = - \log(k_R^+/k_R^-),
\end{equation}
must hold, where $\beta\Delta\epsilon_R$ is inferred from the master curve fit
to fold-change measurements, and $k_R^+$ and $k_R^-$ are inferred in some
orthogonal manner. Single molecule measurements such as from~\cite{Hammar2014}
have directly observed these rates, and in the remainder of this work we explore
a complementary approach: inferring repressor binding and unbinding rates
$k_R^+$ and $k_R^-$ from FISH measurements of mRNA population distributions.