### 2020-05-21 doc/appendix_02_constitutive_promoters.tex

% !TEX root = ./busty_transcription.tex
\section{Derivations for non-bursty promoter models}
\subsection{Nonequilibrium Model One - Poisson Promoter}
\mmnote{Derive mean mRNA referred to in means section.}

\subsection{Nonequilibrium Model Two - RNAP Bound and Unbound States}
\mmnote{Obtain mean mRNA and Fano factor from CME.}

\subsection{Nonequilibrium Model Three
    - Multistep Transcription Initiation and Escape}
\mmnote{Derive mean mRNA from CME.}
\newline
\mmnote{Obtain Fano factor by taking limit from Mitarai et al's expression.}

\subsection{Nonequilibrium Model Four
    - ``Active'' and ``Inactive'' States}
\mmnote{Derive mean mRNA from CME.}
\newline
\mmnote{Obtain Fano factor by relabeling cartoon from Jones2014.}