### 2020-05-26 doc/abstract.tex
RP work on May 26
I worked on the Abstract, the Intro, the Discussion and put in a label in Section 4 so when it was called the section got the right number.

% !TEX root = ./busty_transcription.tex
\begin{abstract}
The study of transcription remains one of the centerpieces of modern biology
with implications in settings from development to metabolism to evolution to
disease. Precision measurements using a host of different techniques including
fluorescence and sequencing readouts have raised the bar for what it means to
quantitatively understand transcriptional regulation. In particular our
understanding of the simplest genetic circuit is sufficiently refined both
experimentally and theoretically that it has become possible to carefully
discriminate between different conceptual pictures of how this regulatory system
works. This regulatory motif, originally posited by Jacob and Monod in the
1960s, consists of a single transcriptional repressor binding to a promoter site
and inhibiting transcription. In this paper, we show how seven distinct models
of this so-called simple-repression motif, based both on equilibrium and kinetic
thinking, can be used to derive the predicted levels of gene expression. These
different models are invoked to confront a variety of different data on mean,
variance and full gene expression distribution, illustrating the extent to which
such models can be distinguished, and suggesting a two-state model with a distribution
of burst sizes as the most potent of the seven for describing the simple-repression motif.
\end{abstract}