### 2020-06-02 doc/section_05_discussion.tex
RP June 2
Intro has references to Muller-Hill
Discussion expanded
bib has many new entries

% !TEX root = ./busty_transcription.tex
\section{Discussion and future work}

The study of gene expression is one of the dominant themes of modern biology,
made all the more urgent by the dizzying pace at which genomes are being
sequenced. But there is a troubling Achilles heel buried in all of that genomic
data and that is our inability to find and interpret regulatory sequence. In
many cases, this is not possible even qualitatively, let alone the possibility
of quantitative dissection of the regulatory parts of genomes in a predictive
fashion. Other recent work has tackled the challenge of finding and annotating
the regulatory part of genomes \cite{Belliveau2018, Ireland2020}. Once we have
determined the architecture of the regulatory part of the genome, we are then
faced with the next class of questions which are sharpened by formulating them
in mathematical terms, namely, what are the input-output properties of these
regulatory circuits and what knobs control them?

The present work has tackled that question in the context of the first
regulatory architecture hypothesized in the molecular biology era, namely, the
repressor-operator model of Jacob and Monod~\cite{Jacob1961}. Regulation in that
architecture is the result of a competition between a repressor which inhibits
transcription and RNAP polymerase which undertakes it. Through the labors of
generations of geneticists, molecular biologists and biochemists, an
overwhelming amount of information and insight has been garnered into this
simple regulatory motif, licensing it as what one might call the ``hydrogen
atom'' of regulatory biology. It is from that perspective that the present paper
explores the extent to which some of the different models that have been
articulated to describe that motif allow us to understand both the average level
of gene expression found in a population of cells, the intrinsic cell-to-cell
variability, and the full gene expression distribution found in such population
as would be reported in a single molecule Fluorescence in situ Hybridization
experiment, for example.

Our key insights can be summarized as follows. First, as shown in
Figure~\ref{fig1:means_cartoons}, the mean expression in the simple repression
architecture is captured by a master curve in which the action of repressor and
the details of the polymerase interaction with the promoter appear separately
and additively in an effective free energy. Interestingly, as has been shown
elsewhere in the context of the Monod-Wyman-Changeux model, these kinds of
coarse-graining results are an exact mathematical result and do not constitute
hopeful approximations or biological naivete \cite{Razo-Mejia2018, Chure2019}.
To further dissect the relative merits of the different models, we must appeal
to higher moments of the gene expression probability distribution. To that end,
our second set of insights focus on gene expression noise, where it is seen that
a treatment of the constitutive promoter already reveals that some models have
Fano factors (variance/mean) that are less than one, at odds with any and all
experimental data that we are aware of\cite{So2011, Jones2014}. This theoretical
result allows us to directly discard a subset of the models (models 1-3 in
Figure~\ref{fig2:constit_cartoons}(A)) since they cannot be reconciled with
experimental observations. The two remaining models (models 4 and 5 in
Figure~\ref{fig2:constit_cartoons}) appear to contain enough microscopic realism
to be able to reproduce the data. A previous exploration of model 4 demonstrated
the sloppy nature of the model in which data on single-cell mRNA counts alone
cannot constrain the value of all parameters simultaneously
\cite{Razo-Mejia2020}. Here we demonstrate that the proposed one-state bursty
promoter model (model 5 in Figure~\ref{fig2:constit_cartoons}) emerges as  a
limit of the commonly used two-state promoter model \cite{Peccoud1995,
Shahrezaei2008, So2011, Sanchez2013, Jones2014}. We put the idea to the test
that this level of coarse-graining is rich enough to reproduce previous
experimental observations. In particular we perform Bayesian inference to
determine the two parameters describing the full steady-state mRNA distribution,
finding that the model is able to provide a quantitative description of a
plethora of promoter sequences with different mean levels of expression and
noise.

With the results of the constitutive promoter in hand, we then fix the
parameters associated with this class of promoters and use them as input for
evaluating the noise in gene expression for the simple repression motif itself.
This allows us to provide a single overarching analysis of both the constitutive
and simple repression architectures using one simple model and corresponding set
of self-consistent parameters, demonstrating not only a predictive framework,
but also reconciling the equilibrium and non-equilibrium views of the same
simple repression constructs. More specifically, we obtained values for the
transcription factor association and dissociation rates by performing Bayesian
inference on the full mRNA distribution for data obtained from simple-repression
promoters with varying number of transcription factors per cell and affinity of
such transcription factors for the binding site. The free energy value obtained
from these kinetic rates -- computed as the log ratio of the rates -- agrees
with previous inferences performed only from mean gene expression measurements,
that assumed an equilibrium rather than a kinetic
framework~\cite{Garcia2011a,Razo-Mejia2018}.

It is interesting to speculate what microscopic details are being coarse-grained
by our burst rate and burst size in Figure~\ref{fig2:constit_cartoons} model 5.
Chromosomal locus is one possible influence we have not addressed in this work,
as all the single-molecule mRNA data from~\cite{Jones2014} that we considered
was from a construct integrated at the \textit{galK} locus. The results
of~\cite{Chong2014} indicate that transcription-induced supercoiling contributes
substantially in driving transcriptional bursting, and furthermore, their
Figure~7 suggests that the parameters describing the rate, duration, and size of
bursts vary substantially for transcription from different genomic loci.
Although the authors of~\cite{Englaender2017} do not address noise, they note
enormous differences in mean expression levels when an identical construct is
integrated at different genomic loci. The authors of~\cite{Engl2020} attribute
noise and burstiness in their single-molecule mRNA data to the influence of
different sigma factors, which is a reasonable conclusion from their data. Could
the difference also be due to the different chromosomal locations of the two
operons?

It seems clear therefore that chromosomal locus plays a major role in mean
expression levels as well as noise. What features of different loci are and are
not important? Sigma factors could certainly play a part, but it seems many
other mechanisms are at play: supercoiling, and more generally 3D organization
of the chromosome, read-through of neighboring genes, and surely some
fascinating unknown unknowns. Could our preferred coarse-grained model capture
the variability across different loci? If so, and we were to repeat the
parameter inference as done in this work, is there a simple theoretical model we
could build to understand the resulting parameters?

\textcolor{red}{RP: I relegated all remaining comments to the background.  Muir
can look at them to see if he really wants them.  I agreed with Manuel that the stuff
below should not be included in the Discussion.  ONE THING FOR SURE - Muller-Hill
and Oehler need to get a lot of love in references at beginning.}
In summary, this work took up the challenge of exploring the extent
to which a single specific mechanistic model of the simple-repression regulatory
architecture suffices to explain the broad sweep of experimental
data for this system.  Pioneering early experimental efforts from
the M\"{u}ller-Hill lab established the simple-repression motif
as an arena for the quantitative dissection of regulatory response in
bacteria, with similar beautiful work emerging in examples such as the
arabinose and gal operons as well~\cite{Dunn1984b, Oehler1990, Oehler1994, Schleif2000,Weickert1993, Semsey2002, SwintKruse2009}.  In light of a new generation of precision measurements
on these systems, the definition of what it means to understand them can now
be formulated as a rigorous quantitative question, and many aspects of
the simple repression motif seem to merit the claim that they are sufficiently
well understood to use that understanding to design new versions of that
architecture based upon predictions about how repressor copy number and DNA binding site strength
control expression.  In our view, the next step in the progression is
to first perform similar rigorous analyses of the fundamental ``basis set'' of regulatory
architectures such as the simple activation motif that is regulated by
a single activator binding site, the repressor-activator architecture, mediated by
the binding of both a single activator and a single repressor and beyond.
With the individual input-output functions in hand, similar quantitative
dissections including the rigorous analysis of their tuning parameters can be undertaken
for the ``basis set'' of full gene-regulatory networks such as switches, feed-forward architectures and
oscillators for example, building upon the recent impressive bonanza of efforts from
systems biologists and synthetic biologists~\cite{Milo2002,Alon2007}.

%\mrm{We need to decide ASAP if this is going to be included in the discussion or not. My vote is no, but Jupiter must decide.}
%\mmnote{Do all my approximations work simply because galK happens to have a low
%duty cycle of supercoiling, $\beta/\alpha$ in Chong2014 picture? A potentially
%interesting test: put in the dozen or so Hernan constructs at some other locus,
%preferably one of the ones Chong identified as having a large duty cycle (large
%$\beta/\alpha$), i.e., the opposite limit of galK. Then just measure
%fold-changes. You don't even need to do FISH, although that would be even
%better. If their model is right, $\rho$ should be approximately set by the duty
%cycle of supercoiling, NOT by RNAP kinetics.
%\textit{This is a NICE prediction for future work.} \textit{cspE}
%is claimed nonessential, prediction is you drop HG reporter constructs in there
%and the inferred binding energies of all lacI operators shift the same amount,
%by about $\ln(3)\sim 1kT$. Not huge, but detectable, although might be
%confounding effects from moving to a new locus. With FISH measurements, $k^+$
%and $k^-$ in Fig 2, model 4 should now be separately inferrable if Chong2014
%measurements of $\beta/\alpha$ are correct. This would require an obnoxious
%amount of cloning, but once you did that, the measurements would be same old
%stuff that RP lab does.}
%
%A direct \mmnote{if cloning intensive} future experiment would be
%to design a payload with a promoter driving a FISH reporter,
%integrate that payload at a variety of loci scattered around the genome,
%and carry out FISH measurements of the resulting strains.
%\mmnote{
%Did Brewster already do this? I remember he integrated arrays of
%competitor binding sites all over the genome, but did he put
%lacUV5 + reporter scattered around?}
%Obviously integration loci would need to be carefully chosen so
%that one could plausibly hope the deletion and integration did
%not disturb the cell's physiology too badly.
%
%\mrm{Rather than proposing more experiments, I feel that we are missing a closing paragraph that puts all these findings in context. Where do we go from here as a community? Things such as the next building block, i.e. simple activation, and then from there scale up the complexity until we can get a full gene regulatory network kind of thing. Rather than going small into tiny details specific to this, let's step back and look at the big picture.}