### 2020-06-14 doc/section_07_acknowledgments.tex
add missing sections

% !TEX root = ./busty_transcription.tex
\section*{Acknowledgments}

We thank Rob Brewster for providing the raw single-molecule mRNA FISH data. We
thank Justin Bois for his key support with the Bayesian inference section. We
would also like to thank Griffin Chure for invaluable feedback on the
manuscript. This material is based upon work supported by the National Science
Foundation Graduate Research Fellowship under Grant No. DGE‐1745301. This work
was also supported by La Fondation Pierre-Gilles de Gennes, the Rosen Center at
Caltech, and the NIH 1R35 GM118043 (MIRA). M.R.M. was supported by the Caldwell
CEMI fellowship.