### 2020-09-08 doc/resubmission/section_02_means.tex
first attempt to shrink section 02. Sending to appendix 01 in the mean time'

% !TEX root = ./busty_transcription.tex
\section{Mean Gene Expression}\label{section_02_means}

As noted in the previous section, there are two broad classes of models in play
for computing the input-output functions of regulatory architectures as shown in
Figure~\ref{fig1:means_cartoons}. In both classes of model, the promoter is
imagined to exist in a discrete set of states of occupancy, with each such state
of occupancy accorded its own rate of transcription -- including no 
transcription for many of these states. The models are probabilistic with each
state assigned some probability and the overall rate of transcription given by 
\begin{equation}
\mbox{average rate of transcription} = \sum_i r_i p_i,
\label{eq:transcrip_prop_pbound}
\end{equation}
where $i$ labels the distinct states, $p_i$ is the probability of the
$i^{\text{th}}$ state, and $r_i$ is the rate of transcription of that state.
Ultimately, the different models differ along several key axes: what states to
consider and how to compute the probabilities of those states.

The first class of models that are the focus of the present section on
predicting the mean level of gene expression, sometimes known as thermodynamic
models, invoke the tools of equilibrium statistical mechanics to compute the
probabilities of the promoter microstates~\cite{Ackers1982, Shea1985,
Buchler2003, Vilar2003a, Vilar2003b, Bintu2005a, Bintu2005c, Gertz2009,
Sherman2012, Saiz2013}. As seen in Figure~\ref{fig1:means_cartoons}(B), even
within the class of thermodynamic models, we can make different commitments
about the underlying microscopic states of the promoter.  Indeed, the list of
options considered here does not at all exhaust the suite of different
microscopic states we can assign to the promoter.

The second class of models that allow us to access the mean gene expression use
chemical master equations to compute the probabilities of the different
microscopic states ~\cite{Ko1991, Peccoud1995, Record1996, Kepler2001,
Sanchez2008, Shahrezaei2008, Sanchez2011, Michel2010}. As seen in
Figure~\ref{fig1:means_cartoons}(C), we consider a host of different
nonequilibrium models, each of which will have its own result for both the mean
(this section) and noise (next section) in gene expression.

\subsection{Fold-changes are indistinguishable across models}
As a first stop on our search for the ``right'' model of simple repression, let
us consider what we can learn from theory and experimental measurements on the
average level of gene expression in a population of cells. One experimental
strategy that has been particularly useful (if incomplete since it misses out on
gene expression dynamics) is to measure the fold-change in mean
expression~\cite{Garcia2011}. The fold-change is defined as
\begin{equation}
\text{fold-change}
= \frac{\langle \text{gene expression with repressor present} \rangle}
        {\langle \text{gene expression with repressor absent} \rangle}
= \frac{\langle m (R > 0) \rangle}{\langle m (R = 0) \rangle}
\label{eq:fc_def}
\end{equation}
where angle brackets $\left\langle \cdot \right\rangle$ denote the average over
a population of cells and mean mRNA $\langle m\rangle$ is viewed as a function
of repressor copy number $R$. What this means is that the fold-change in gene
expression is a relative measurement of the effect of the transcriptional
repressor ($R > 0$) on the gene expression level compared to an unregulated
promoter ($R = 0$). The second equality in Eq.~\ref{eq:fc_def} follows from
assuming that the translation efficiency, i.e., the number of proteins
translated per mRNA, is the same in both conditions. In other words, we assume
that mean protein level is proportional to mean mRNA level, and that the
proportionality constant is the same in both conditions and therefore cancels
out in the ratio. This is reasonable since the cells in the two conditions are
identical except for the presence of the transcription factor, and the model
assumes that the transcription factor has no direct effect on translation.

Fold-change has proven a very convenient observable in past
work~\cite{Garcia2011a, Brewster2014, Razo-Mejia2018, Chure2019}. Part of its
utility in dissecting transcriptional regulation is its ratiometric nature,
which removes many secondary effects that are present when making an absolute
gene expression measurement. Also, by measuring otherwise identical cells with
and without a transcription factor present, any biological noise common to both
conditions can be made to cancel away.

\fig{fig1:means_cartoons} depicts a smorgasbord of mathematicized cartoons for
simple repression using both thermodynamic and kinetic models that have appeared
in previous literature. For each cartoon, we calculate the fold-change in mean
gene expression as predicted by that model, deferring some algebraic details to
Appendix~\ref{sec:non_bursty}. What we will find is that all cartoons collapse
to a single master curve, shown in \fig{fig1:means_cartoons}(D), which contains
just two parameters. We label the parameters $\Delta F_R$, an effective free
energy parametrizing the repressor-DNA interaction, and $\rho$, which subsumes
all details of transcription in the absence of repressors. We will offer some
intuition for why this master curve exists and discuss why at the level of the
mean expression, we are unable to discriminate ``right'' from ``wrong'' cartoons
given only measurements of fold-changes in expression.

\subsubsection{Two- and Three-state Thermodynamic Models}
We begin our analysis with models (1) and (2) in
Figure~\ref{fig1:means_cartoons}(B). In each of these models the promoter
is idealized as existing in a set of discrete states; the difference being
whether or not the RNAP bound state is included or not. Gene expression is then
assumed to be proportional to the probability of the promoter being in either
the empty state (model (1)) or the RNAP-bound state (model (2)). We direct the
reader to Appendix~\ref{sec:non_bursty} for details on the derivation of the
fold-change. For our purposes here, it suffices to state that the functional
form of the fold-change for model (1) is
\begin{equation}
\text{fold-change}
= \left(1 + \frac{R}{N_{NS}} e^{-\beta\Delta\varepsilon_R}\right)^{-1},
\end{equation}
where $R$ is the number of repressors per cell, $N_{NS}$ is the number of
non-specific binding sites where the repressor can bind, $\Delta\varepsilon_R$
is the repressor-operator binding energy, and $\beta \equiv (k_BT)^{-1}$. This
equation matches the form of the master curve in
Figure~\ref{fig1:means_cartoons}(D) with $\rho=1$ and $\Delta F_R =
\beta\Delta\varepsilon_r - \log (R / N_{NS})$. For model (2) we have a similar
situation. The fold-change takes the form
\begin{align}
\text{fold-change}
&= \left(
1 + \frac{\frac{R}{N_{NS}} e^{-\beta\Delta\varepsilon_R}}
        {1 + \frac{P}{N_{NS}} e^{-\beta\Delta\varepsilon_P}}
\right)^{-1}
\\
&= (1 + \exp(-\Delta F_R + \log\rho))^{-1},
\end{align}
where $P$ is the number of RNAP per cell, and $\Delta\varepsilon_P$ is the
RNAP-promoter binding energy. For this model we have $\Delta F_R =
\beta\Delta\varepsilon_R - \log(R/N_{NS})$ and $\rho = 1 +
\frac{P}{N_{NS}}\mathrm{e}^{-\beta\Delta\varepsilon_P}$. Thus far, we see that
the two thermodynamic models, despite making different coarse-graining
commitments, result in the same functional form for the fold-change in mean gene
expression.  We now explore how kinetic models fare when faced with computing
the same observable.

\subsubsection{Kinetic models}
One of the main difference between models shown in
Figure~\ref{fig1:means_cartoons}(C), casted in the language of chemical master
equations, compared with the thermodynamic models discussed in the previous 
section is the probability space over which they are built. Rather than keeping
track only of the microstate of the promoter, and assuming that gene expression
is proportional to the probability of the promoter being in a certain
microstate, chemical master equation models are built on the entire probability
state of both the promoter microstate, and the current mRNA count. Therefore, in
order to compute the fold-change, we must compute the mean mRNA count on each of
the promoter microstates, and add them all together~\cite{Sanchez2013}.

Again, we spare all details of the derivation to Appendix~\ref{sec:non_bursty}.
Here we just highlight the general findings for all five kinetic models. As
already shown in Figure~\ref{fig1:means_cartoons}(C) and (D), all the kinetic
models explored can be collapsed into the master curve. Given that the
repressor-bound state only connects to the rest of the promoter dynamics via its
binding and unbinding rates, $k_R^+$ and $k_R^-$ respectively, all models can
effectively be separated into two categories: a single repressor-bound state,
and all other promoter states with different levels of coarse graining. This
structure then guarantees that, at steady-state, detailed balance between these
two groups is satisfied. What this implies is that the steady-state distribution
of each of the non-repressor states has the same functional form with or without
the repressor, allowing us to write the fold-change as a product of the ratio of
the binding and unbinding rates of the promoter, and the promoter details. This
is 
\mrm{This is based in one of the comments from reviewer \# 2. It is empirically
true for all models we explored, but I don't know how to prove that it is a 
general statement for any unimaginable complex model.}
\begin{align}
\text{fold-change} &= \left( 1 + \frac{k_R^+}{k_R^-} \rho \right)^{-1},\\
&= (1 + \exp(-\Delta F_R + \log(\rho) ))^{-1},
\end{align}
where $\Delta F_R \equiv -\log(k_R^+/k_R^-)$, and the functional forms of $\rho$
for each model change as shown in Figure~\ref{fig1:means_cartoons}(C) (See
Appendix~\ref{sec:non_bursty} for further details).

The key outcome of our analysis of the models in
Figure~\ref{fig1:means_cartoons} is the existence of a master curve shown
in Figure~\ref{fig1:means_cartoons}(D) to which the fold-change predictions
of all the models collapse. This master curve is parametrized by only two
effective parameters: $\Delta F_R$, which characterizes the number of repressors
and their binding strength to the DNA, and $\rho$, which characterizes all other
features of the promoter architecture. The key assumption underpinning this
result is that no transcription occurs when a repressor is bound to its
operator. Given this outcome, a mean-based metric such as the fold-change that
can be readily measured experimentally is insufficient to discern between these
different levels of coarse-graining. In the next section we extend the analysis
of the models to higher moments of the mRNA distribution as we continue to look
for the ``right'' level of coarse graining.

% \subsection{Discussion of Results Across Models for Fold-Changes in Mean
% Expression}
% The key outcome of our analysis of the models in
% Figure~\ref{fig1:means_cartoons} is the existence of a master curve shown
% in Figure~\ref{fig1:means_cartoons}(D) to which the fold-change predictions
% of all the models collapse. This master curve is parametrized by only two
% effective parameters: $\Delta F_R$, which characterizes the number of repressors
% and their binding strength to the DNA, and $\rho$, which characterizes all other
% features of the promoter architecture. The key assumption underpinning this
% result is that no transcription occurs when a repressor is bound to its
% operator. Note, however, that we are agnostic about the molecular mechanism
% which achieves this; steric effects are one plausible mechanism, but, for
% instance, ``action-at-a-distance'' mediated by kinked DNA due to repressors
% bound tens or hundreds of nucleotides upstream of a promoter is plausible as
% well.

% Why does the master curve of Figure~\ref{fig1:means_cartoons}(D) exist at all?
% This brings to mind the deep questions posed in,
% e.g.,~\cite{Frank2013} and~\cite{Frank2014a}, suggesting we consider multiple
% plausible models of a system and search for their common patterns to tease
% out which broad features are and are not important.
% In our case, the key feature seems to be the
% exclusive nature of repressor and RNAP binding, which allows the parameter
% describing the repressor, $\Delta F_R$, to cleanly separate from all other
% details of the promoter architecture, which are encapsulated in $\rho$.
% Arbitrary nonequilibrium behavior can occur on the rest of the
% promoter state space, but it may all be swept up in the effective
% parameter $\rho$, to which the repressor makes no contribution.
% We point the interested reader to~\cite{Gunawardena2012}
% and~\cite{Ahsendorf2014} for an interesting analysis of similar problems
% using a graph-theoretic language.

% As suggested in~\cite{Chure2019}, we believe this master curve should generalize
% to architectures with multiple repressor binding sites, as long as the
% exclusivity of transcription factor binding and transcription initiation is
% maintained. The interpretation of $\Delta F_R$ is then of an effective free
% energy of all repressor bound states. In an equilibrium picture this is simply
% given by the log of the sum of Boltzmann weights of all repressor bound states,
% which looks like the log of a partition function of a subsystem. In a
% nonequilibrium picture, while we can still mathematically gather terms and give
% the resulting collection the label $\Delta F_R$, it is unclear if the physical
% interpretation as an effective free energy makes sense. The problem is that free
% energies cannot be assigned unambiguously to states out of equilibrium because
% the free energy change along a generic path traversing the state space is path
% dependent, unlike at equilibrium. A consequence of this is that, out of
% equilibrium, $\Delta F_R$ is no longer a simple sum of Boltzmann weights.
% Instead it resembles a restricted sum of King-Altman diagrams~\cite{King1956,
% Hill1966}. Following the work of Hill~\cite{Hill1989}, it may yet be possible to
% interpret this expression as an effective free energy, but this remains unclear
% to us. We leave this an open problem for future work.

% If we relax the requirement of exclusive repressor-RNAP binding, one could
% imagine models in which repressor and RNAP doubly-bound states are allowed,
% where the repressor's effect is to \text{reduce} the transcription rate rather
% than setting it to zero. Our results do not strictly apply to such a model,
% although we note that if the repressor's reduction of the transcription rate is
% substantial, such a model might still be well-approximated by one of the models
% in~\fig{fig1:means_cartoons}.

% One may worry that our ``one curve to rule them all'' is a mathematical
% tautology. In fact we \textit{agree} with this criticism if $\Delta F_R$ is
% ``just a fitting parameter'' and cannot be meaningfully interpreted as a real,
% physical free energy. An analogy to Hill functions is appropriate here. One of
% their great strengths and weaknesses, depending on the use they are put to, is
% that their parameters coarse-grain many details and are generally not
% interpretable in terms of microscopic models, for deep reasons discussed at
% length in~\cite{Frank2013}. By contrast, our master curve claims to have the
% best of both worlds: a coarse-graining of all details besides the repressor into
% a single effective parameter $\rho$, while simultaneously retaining an
% interpretation of $\Delta F_R$ as a physically meaningful and interpretable free
% energy. Our task, then, is to prove or disprove this claim.

% How do we test this and probe the theory with fold-change measurements? There is
% a fundamental limitation in that the master curve is essentially a one-parameter
% function of $\Delta F_R + \log\rho$. Worse, there are many \textit{a priori}
% plausible microscopic mechanisms that could contribute to the value of $\rho$,
% such as RNAP binding and escape kinetics~\cite{DeHaseth1998, Mitarai2015},
% and/or supercoiling accumulation and release~\cite{Chong2014, Sevier2016},
% and/or, RNAP clusters analogous to those observed in
% eukaryotes~\cite{Cisse2013, Cho2016} and recently also observed in
% bacteria~\cite{Ladouceur2020}. Even if $\Delta F_R$ is measured to high
% precision, inferring the potential microscopic contributions to $\rho$, buried
% inside a log no less, from fold-change measurements seems beyond reach. As a
% statistical inference problem it is entirely nonidentifiable, in the language
% of~\cite{Gelman2013}, Section 4.3.

% If we cannot simply infer values of $\rho$ from measurements of fold-change, can
% we perturb some of the parameters that make up $\rho$ and measure the change?
% Unfortunately we suspect this is off-limits experimentally: most of the
% potential contributors to $\rho$ are global processes that affect many or all
% genes. For instance, changing RNAP association rates by changing RNAP copy
% numbers, or changing supercoiling kinetics by changing topoisomerase copy
% numbers, would massively perturb the entire cell's physiology and confound any
% determination of $\rho$.

% One might instead imagine a bottom-up modeling approach, where we mathematicize
% a model of what we hypothesize the important steps are and are not, use \textit{in vitro}
% data for the steps deemed important, and \textit{predict} what $\rho$ should be.
% But again, because of the one-parameter nature of the master curve, many
% different models will likely make indistinguishable predictions, and without any
% way to experimentally perturb \textit{in vivo}, there is no clear way
% to test whether the modeling assumptions are correct.

% In light of this, we prefer the view that parameters and rates are not directly
% comparable between cartoons in~\fig{fig1:means_cartoons}. Rather, parameters in
% the simpler cartoons represent coarse-grained combinations of parameters in the
% finer-grained models. For instance, by equating $\rho$ between any two models,
% one can derive various possible correspondences between the two models'
% parameters. Note that these correspondences are clearly not unique, since many
% possible associations could be made. It then is a choice as to what microscopic
% interpretations the model-builder prefers for the parameters in a particular
% cartoon, and as to which coarse-grainings lend intuition and which seem
% nonsensical. Indeed, since it remains an open question what microscopic features
% dominate $\rho$ (as suggested above, perhaps RNAP binding and escape
% kinetics~\cite{DeHaseth1998, Mitarai2015}, or supercoiling accumulation and
% release~\cite{Chong2014, Sevier2016}, or, something more exotic like RNAP
% clusters~\cite{Cisse2013, Cho2016, Ladouceur2020}), we are hesitant to put too
% much weight on any one microscopic interpretation of model parameters that make
% up $\rho$.

% One possible tuning knob to probe $\rho$ that would not globally perturb the
% cell's physiology is to manipulate RNAP binding sites. Work such
% as~\cite{Kinney2010} has shown that models of sequence-dependent RNAP affinity can be
% inferred from data, and the authors of~\cite{Brewster2012} showed that the model
% of~\cite{Kinney2010} has predictive power by using the model to \textit{design}
% binding sites of a desired affinity. But for our purposes, this begs the
% question: the authors of~\cite{Kinney2010} \textit{assumed} a particular model
% (essentially our 3-state equilibrium model but without the repressor), so it is
% unclear how or if such methods can be turned around to \textit{compare}
% different models of promoter function.

% Another possible route to dissect transcription details without a global
% perturbation would be to use phage polymerase with phage-specific promoters.
% While such results would carry some caveats, e.g., whether the repression of
% the phage polymerase is a good analog to the repression of the native RNAP,
% it could nevertheless be worthy of consideration.

% We have already pointed out that the master curve of~\fig{fig1:means_cartoons}
% is essentially a one-parameter model, the one parameter being $\Delta F_R +
% \log\rho$. By now the reader may be alarmed as to how can we even determine
% $\Delta F_R$ and $\rho$ independently of each other, never mind shedding a lens
% on the internal structure of $\rho$ itself. A hint is provided by the weak
% promoter approximation, invoked repeatedly in prior studies~\cite{Bintu2005c,
% Garcia2011a, Razo-Mejia2018} of simple repression using the 3-state equilibrium
% model in~\fig{fig1:means_cartoons}(B). In that picture, the weak promoter
% approximation means $\frac{P}{N_{NS}}\exp(-\beta\Delta\varepsilon_P) \ll 1$,
% meaning therefore $\rho\approx1$.  This approximation can be well justified on
% the basis of the number of RNAP and $\sigma$ factors per cell
% and the strength of binding of RNAP to
% DNA at weak promoters. This is suggestive, but how can we be sure that $\rho$ is
% not, for instance, actually $10^2$ and that $\Delta F_R$ hides a compensatory
% factor? A resolution is offered by an independent inference of $\rho$ in the
% absence of repressors. This was done in~\cite{Razo-Mejia2020} by fitting
% nonequilibrium model 4 in~\fig{fig1:means_cartoons}(C), with zero repressor
% (looking ahead, this is equivalent to model 4
% in Figure~\ref{fig2:constit_cartoons}(A)),
% to single-cell mRNA counts data from~\cite{Brewster2014}. This provided a
% determination of $k^+$ and $k^-$, from which their ratio is estimated to be no
% more than a few $10^{-1}$ and possibly as small as $10^{-2}$.

% The realization that $\rho\approx1$ to an excellent approximation,
% \textit{independent} of which model in~\fig{fig1:means_cartoons} one prefers,
% goes a long way towards explaining the surprising success of equilibrium models
% of simple repression. Even though our 2- and 3-state models get so many details
% of transcription wrong, it does not matter because fold-change is a cleverly
% designed ratio. Since $\rho$ subsumes all details except the repressor, and
% $\log\rho\approx0$, fitting these simple models to fold-change measurements can
% still give a surprisingly good estimate of repressor binding energies. So the
% ratiometric construction of fold-change fulfills its intended purpose of
% canceling out all features of the promoter architecture except the repressor
% itself. Nevertheless it is perhaps surprising how effectively it does so:
% \textit{a priori}, one might not have expected $\rho$ to be quite so close to 1.

% We would also like to highlight the relevance of~\cite{Landman2019} here.
% Landman et.\ al.\ reanalyzed and compared \textit{in vivo} and
% \textit{in vitro} data on the lacI repressor's binding affinity to its
% various operator sequences. (The \textit{in vivo}
% data was from, essentially, fitting our master curve to expression
% measurements.) They find broad agreement between the
% \textit{in vitro} and \textit{in vivo}
% values. This reinforces the suspicion that the equilibrium $\Delta\varepsilon_R$
% repressor binding energies do in fact represent real physical free energies.
% Again, \textit{a priori} this did not have to be the case, even knowing that
% $\rho\approx1$.

% In principle, if $\Delta F_R$ can be measured to sufficient precision, then
% deviations from $\rho=1$ become a testable matter of experiment. In practice, it
% is probably unrealistic to measure repressor rates $k_R^+$ or $k_R^-$ or
% fold-changes in expression levels (and hence $\Delta\varepsilon_R$) precisely
% enough to detect the expected tiny deviations from $\rho=1$. We can estimate the
% requisite precision in $\Delta F_R$ to resolve a given $\Delta\rho$ by noting,
% since $\rho\approx1$, that $\log(1+\Delta\rho)\approx \Delta\rho$, so
% $\Delta(\Delta F_R) \approx \Delta\rho$. Suppose we are lucky and $\Delta\rho$
% is $\sim0.1$, on the high end of our range estimated above. A
% determination of $\Delta\varepsilon_R/k_BT$ with an uncertainty of barely 0.1
% was achieved in the excellent measurements of~\cite{Razo-Mejia2018}, so this
% requires a very difficult determination of $\Delta F_R$ for a very crude
% determination of $\rho$, which suggests, to put it lightly, this is not a
% promising path to pursue experimentally. It is doubtful that inference of
% repressor kinetic rates would be any easier.%\footnote{\mmnote{In fact the
% % uncertainties quoted in Hammar et.\ al.~\cite{Hammar2014} from Elf's group are
% % small enough to do this calculation, but HJ \& I think their error bars are
% % insanely optimistic. Not sure I want to blatantly call them out on it though??}}

% Moving forward, we have weak evidence supporting the interpretation of $\Delta
% F_R$ as a physically real free energy~\cite{Landman2019} and other work casting
% doubt~\cite{Hammar2014}. How might we resolve the confusion? If there is no
% discriminatory power to test the theory and distinguish the various models with
% measurements of fold-changes in means, how do we probe the theory? Clearly to
% discriminate between the nonequilibrium models in~\fig{fig1:means_cartoons}, we
% need to go beyond means to ask questions about kinetics, noise and even full
% distributions of mRNA copy numbers over a population of cells. If the
% ``one-curve-to-rule-them-all'' is more than a mathematical tautology, then the
% free energy of repressor binding inferred from fold-change measurements should
% agree with repressor binding and unbinding rates. In other words, the
% equilibrium and nonequilibrium definitions of $\Delta F_R$ must agree, meaning
% \begin{equation}
% \Delta F_R = \beta\Delta\varepsilon_R - \log(R/N_{NS})
%         = - \log(k_R^+/k_R^-),
% \label{eq:deltaFR_eq_noneq_equiv}
% \end{equation}
% must hold, where $\beta\Delta\varepsilon_R$ is inferred from the master curve
% fit to fold-change measurements, and $k_R^+$ and $k_R^-$ are inferred in some
% orthogonal manner. Single molecule measurements such as from~\cite{Hammar2014}
% have directly observed these rates, and in the remainder of this work we explore
% a complementary approach: inferring repressor binding and unbinding rates
% $k_R^+$ and $k_R^-$ from single-molecule measurements of mRNA population
% distributions.