### 2020-09-24 doc/resubmission/section_05_discussion.tex

% !TEX root = ./busty_transcription.tex
\section{Discussion and future work}

The study of gene expression is one of the dominant themes of modern biology,
made all the more urgent by the dizzying pace at which genomes are being
sequenced. But there is a troubling Achilles heel buried in all of that genomic
data, which is our inability to find and interpret regulatory sequence. In
many cases, this is not possible even qualitatively, let alone the possibility
of quantitative dissection of the regulatory parts of genomes in a predictive
fashion. Other recent work has tackled the challenge of finding and annotating
the regulatory part of genomes \cite{Belliveau2018, Ireland2020}. Once we have
determined the architecture of the regulatory part of the genome, we are then
faced with the next class of questions which are sharpened by formulating them
in mathematical terms, namely, what are the input-output properties of these
regulatory circuits and what knobs control them?

The present work has tackled that question in the context of the first
regulatory architecture hypothesized in the molecular biology era, namely, the
repressor-operator model of Jacob and Monod~\cite{Jacob1961}. Regulation in that
architecture is the result of a competition between a repressor which inhibits
transcription and RNAP polymerase which undertakes it. Through the labors of
generations of geneticists, molecular biologists and biochemists, an
overwhelming amount of information and insight has been garnered into this
simple regulatory motif, licensing it as what one might call the ``hydrogen
atom'' of regulatory biology. It is from that perspective that the present paper
explores the extent to which some of the different models that have been
articulated to describe that motif allow us to understand both the average level
of gene expression found in a population of cells, the intrinsic cell-to-cell
variability, and the full gene expression distribution found in such a population
as would be reported in a single molecule mRNA Fluorescence \textit{in situ}
Hybridization experiment, for example.

Our key insights can be summarized as follows. First, as shown in
Figure~\ref{fig1:means_cartoons}, the mean expression in the simple repression
architecture is captured by a master curve in which the action of repressor and
the details of the RNAP interaction with the promoter appear separately and
additively in an effective free energy. Interestingly, as has been shown
elsewhere in the context of the Monod-Wyman-Changeux model, these kinds of
coarse-graining results are an exact mathematical result and do not constitute
hopeful approximations or biological naivete \cite{Razo-Mejia2018, Chure2019}.
The fact that thermodynamic and kinetic models for the simple repression motif
can not only be reconciled, but are indistinguishable from each other at the
level of mean mRNA count is a consequence of the possible transitions between
promoter microstates. The promoter microstates in all models shown in
Figure~\ref{figure1:means_cartoons} can explicitly be separated into two groups:
1) promoters with repressor bound, and 2) all other promoters states. The single
path between these two groups of states guarantees detailed-balance between the
groups. We can therefore separate \textit{expression} related transitions, which
by definition must be out of equilibrium, with \textit{regulation} transitions
that can in principle be in equilibrium. This implies that the probability
distribution of the promoter microstates not involving the repressor take the
same functional form, allowing us to write the fold-change as an effective free
energy involving a regulation term, $\Delta F_R$, and a promoter details term,
$\rho$. We offer this argument as a conjecture, and we suspect that a careful
argument using the graph-theoretic framework based on the Matrix-Tree
Theorem~\cite{Gunawardena2012}, might furnish a ``proof'' not with some 
challenges as this theoretical framework applies, at the present, to finite 
graphs.

To further dissect the relative merits of the different models, we must appeal
to higher moments of the gene expression probability distribution. To that end,
our second set of insights focus on gene expression noise, where it is seen that
a treatment of the constitutive promoter already reveals that some models have
Fano factors (variance/mean) that are less than one, at odds with any and all
experimental data that we are aware of~\cite{So2011, Jones2014}. This
theoretical result allows us to directly discard a subset of the models (models
1-3 in Figure~\ref{fig2:constit_cartoons}(A)) since they cannot be reconciled
with experimental observations. The two remaining models (models 4 and 5 in
Figure~\ref{fig2:constit_cartoons}) appear to contain enough microscopic realism
to be able to reproduce the data. A previous exploration of model 4 demonstrated
the ``sloppy''~\cite{Transtrum2015} nature of the model in which data on
single-cell mRNA counts alone cannot constrain the value of all parameters
simultaneously \cite{Razo-Mejia2020}. Here we demonstrate that the proposed
one-state bursty promoter model (model 5 in
Figure~\ref{fig2:constit_cartoons}(A)) emerges as a limit of the commonly used
two-state promoter model \cite{Peccoud1995, Shahrezaei2008, So2011, Sanchez2013,
Jones2014}. We put the idea to the test that this level of coarse-graining is
rich enough to reproduce previous experimental observations. In particular we
perform Bayesian inference to determine the two parameters describing the full
steady-state mRNA distribution, finding that the model is able to provide a
quantitative description of a plethora of promoter sequences with different mean
levels of expression and noise.

With the results of the constitutive promoter in hand, we then fix the
parameters associated with this class of promoters and use them as input for
evaluating the noise in gene expression for the simple repression motif itself.
This allows us to provide a single overarching analysis of both the constitutive
and simple repression architectures using one simple model and corresponding set
of self-consistent parameters, demonstrating not only a predictive framework,
but also reconciling the thermodynamic and kinetic views of the same simple
repression constructs. More specifically, we obtained values for the
transcription factor association and dissociation rates by performing Bayesian
inference on the full mRNA distribution for data obtained from simple-repression
promoters with varying number of transcription factors per cell and affinity of
such transcription factors for the binding site. The free energy value obtained
from these kinetic rates --computed as the log ratio of the rates-- agrees with
previous inferences performed only from mean gene expression measurements, that
assumed a thermodynamic rather than a kinetic
framework~\cite{Garcia2011a,Razo-Mejia2018}. Interestingly, to constraint the
binding and unbinding rates inferred for Figure~\ref{fig4:repressed_post_full},
all of the rates need to be simultaneously fit. This is because a single
combination of operator (repressor binding site) and aTc concentration
(repressor copy number) cannot constraint both rates independently, but only
their ratio (See Appendix~\ref{sec:bayesian} for further details).

It is important to mention the potential problems present in all models,
including model 5 from Figure~\ref{fig2:constit_cartoons}(A). Irreversible
reactions as the ones used in all transcription events lead to physical
paradoxes such as infinite entropy production rates. These are mathematical
conveniences rather than physical reality, leaving the inferred rates for the
transcription burst rate and burst size as an effective rate with lots of
(potentially) reversible processes. Also, the state of the art modeling imagines
the promoter existing in a series of discrete states. This assumption might not
be entirely true as effects such as DNA supercoiling~\cite{Chong2014,Sevier2016}
or DNA looping~\cite{Boedicker2013a} can be considered continuous
variables; although there are formal ways of integrating out such degrees of
freedom. Furthermore, when computing the fold-change in gene expression
(Section~\ref{section_02_means}) we assume that the translation efficiency of
mRNA transcript is equal in cells with and without regulation. There might be
some known biological effects that could limit the applicability of this
assumption such as cooperation between the transcription and translation 
machinery~\cite{Proshkin2010}, or coupling of mRNA transcription and 
degradation~\cite{Chen2015} among other potential effects. Another implicit 
assumption in all of the models discussed here is that the effects due to 
changes in gene copy number during the cell cycle have no effects on the 
expression profile. For moderate growth rates as the ones used 
in~\cite{Jones2014} from where we obtained the data, there are at any point at
most two copies of the gene. On a recent publication, we explored the effects
that this change in gene dosage can have in gene expression, finding that the
effects are much more drastic at the protein level~\cite{Razo-Mejia2020}.

More specifically, for model 5 from Figure~\ref{fig2:constit_cartoons}(A), it is
interesting to speculate what microscopic details are being coarse-grained by
our burst rate and burst size in Figure~\ref{fig2:constit_cartoons}(A), model 5.
Chromosomal locus is one possible influence we have not addressed in this work,
as all the single-molecule mRNA data from~\cite{Jones2014} that we considered
was from a construct integrated at the \textit{galK} locus. The results
of~\cite{Chong2014} indicate that transcription-induced supercoiling contributes
substantially in driving transcriptional bursting, and furthermore, their
Figure~7 suggests that the parameters describing the rate, duration, and size of
bursts vary substantially for transcription from different genomic loci.
Although the authors of~\cite{Englaender2017} do not address noise, they note
enormous differences in mean expression levels when an identical construct is
integrated at different genomic loci. The authors of~\cite{Engl2020} attribute
noise and burstiness in their single-molecule mRNA data to the influence of
different sigma factors, which is a reasonable conclusion from their data. Could
the difference also be due to the different chromosomal locations of the two
operons? What features of different loci are and are not important? Could our
preferred coarse-grained model capture the variability across different loci? If
so, and we were to repeat the parameter inference as done in this work, is there
a simple theoretical model we could build to understand the resulting
parameters?

In summary, this work took up the challenge of exploring the extent to which a
single specific mechanistic model of the simple-repression regulatory
architecture suffices to explain the broad sweep of experimental data for this
system. Pioneering early experimental efforts from the M\"{u}ller-Hill lab
established the simple-repression motif as an arena for the quantitative
dissection of regulatory response in bacteria, with similar work emerging in
examples such as the \textit{ara} and \textit{gal} operons as
well~\cite{Dunn1984b, Oehler1990, Weickert1993, Oehler1994, Schleif2000,
Semsey2002, SwintKruse2009}. In light of a new generation of precision
measurements on these systems, the definition of what it means to understand
them can now be formulated as a rigorous quantitative question. In particular,
we believe understanding of the simple repression motif has advanced
sufficiently that the design of new versions of the architecture is now
possible, based upon predictions about how repressor copy number and DNA binding
site strength control expression. In our view, the next step in the progression
is to first perform similar rigorous analyses of the fundamental ``basis set''
of regulatory architectures. Natural follow-ups to this work are explorations of
motifs such as simple activation that is regulated by a single activator binding
site, and the repressor-activator architecture, mediated by the binding of both
a single activator and a single repressor, and beyond. With the individual
input-output functions in hand, similar quantitative dissections including the
rigorous analysis of their tuning parameters can be undertaken for the ``basis
set'' of full gene-regulatory networks such as switches, feed-forward
architectures and oscillators for example, building upon the recent impressive
efforts from systems biologists and synthetic biologists~\cite{Milo2002,
Alon2007}.