### 2020-09-26 doc/resubmission/section_02_means.tex

% !TEX root = ./busty_transcription.tex
\section{Mean Gene Expression}\label{section_02_means}

As noted in the previous section, there are two broad classes of models in play
for computing the input-output functions of regulatory architectures as shown in
Figure~\ref{fig1:means_cartoons}. In both classes of model, the promoter is
imagined to exist in a discrete set of states of occupancy, with each such state
of occupancy accorded its own rate of transcription -- including no
transcription for many of these states. This discretization of a potentially
continuous number of promoter states (due to effects such as supercoiling of
DNA~\cite{Chong2014, Sevier2016}  or DNA looping~\cite{Boedicker2013a}) is
analogous to how the Monod-Wyman-Changeux model of allostery coarse-grains
continuous molecule conformations into a finite number of
states~\cite{Martins2011}. The models are probabilistic with each state assigned
some probability and the overall rate of transcription given by 
\begin{equation}
\mbox{average rate of transcription} = \sum_i r_i p_i,
\label{eq:transcrip_prop_pbound}
\end{equation}
where $i$ labels the distinct states, $p_i$ is the probability of the
$i^{\text{th}}$ state, and $r_i$ is the rate of transcription of that state.
Ultimately, the different models differ along several key aspects: what states
to consider and how to compute the probabilities of those states.

The first class of models that are the focus of the present section on
predicting the mean level of gene expression, sometimes known as thermodynamic
models, invoke the tools of equilibrium statistical mechanics to compute the
probabilities of the promoter microstates~\cite{Ackers1982, Shea1985,
Buchler2003, Vilar2003a, Vilar2003b, Bintu2005a, Bintu2005c, Gertz2009,
Sherman2012, Saiz2013}. As seen in Figure~\ref{fig1:means_cartoons}(B), even
within the class of thermodynamic models, we can make different commitments
about the underlying microscopic states of the promoter. Model 1 considers only
two states: a state in which a repressor (with copy number $R$) binds to an
operator and a transcriptionally active state. The free energy difference
between the repressor binding the operator, i.e. a specific binding site, and
one of the $N_{NS}$ non-specific sites is given by $\Delta\varepsilon_R$ (given
in $k_BT$ units as $\beta\equiv (k_BT)^{-1}$). Model 2 expands this model to
include an empty promoter where no transcription happens, as well as a state on
which one of the $P$ RNAPs binds to the promoter with binding energy
$\Delta\varepsilon_P$. Indeed, the list of options considered here does not at
all exhaust the suite of different microscopic states we can assign to the
promoter.

The second class of models that allow us to access the mean gene expression use
chemical master equations to compute the probabilities of the different
microscopic states ~\cite{Ko1991, Peccoud1995, Record1996, Kepler2001,
Sanchez2008, Shahrezaei2008, Sanchez2011, Michel2010}. The main differences
between both modeling approaches can be summarized as: 1) Although for both
classes of models the steps involving transcriptional events are assumed to be
strictly irreversible, thermodynamic models force the regulation, i.e., the
control over the expression exerted by the repressor, to be in equilibrium. This
does not need to be the case for kinetic models. 2) Thermodynamic models ignore
the mRNA count from the state of the Markov process, while kinetic models keep
track of both the promoter state and the mRNA count. 3) Finally, thermodynamic
and kinetic models coarse-grain to different degrees the molecular mechanisms
through which RNAP enters the transcriptional event. As seen in
Figure~\ref{fig1:means_cartoons}(C), we consider a host of different kinetic
models, each of which will have its own result for both the mean (this section)
and noise (next section) in gene expression.

\subsection{Fold-changes are indistinguishable across models}
As a first stop on our search for the ``right'' model of simple repression, let
us consider what we can learn from theory and experimental measurements on the
average level of gene expression in a population of cells. One experimental
strategy that has been particularly useful (if incomplete since it misses out on
gene expression dynamics) is to measure the fold-change in mean
expression~\cite{Garcia2011}. The fold-change is defined as
\begin{equation}
\text{fold-change}
= \frac{\langle \text{gene expression with repressor present} \rangle}
        {\langle \text{gene expression with repressor absent} \rangle}
= \frac{\langle m (R > 0) \rangle}{\langle m (R = 0) \rangle}
= \frac{\langle p (R > 0) \rangle}{\langle p (R = 0) \rangle}
\label{eq:fc_def}
\end{equation}
where angle brackets $\left\langle \cdot \right\rangle$ denote the average over
a population of cells and mean mRNA $\langle m\rangle$ and mean protein $\langle
p\rangle$ are viewed as a function of repressor copy number $R$. What this means
is that the fold-change in gene expression is a relative measurement of the
effect of the transcriptional repressor ($R > 0$) on the gene expression level
compared to an unregulated promoter ($R = 0$). The third equality in
Eq.~\ref{eq:fc_def} follows from assuming that the translation efficiency, i.e.,
the number of proteins translated per mRNA, is the same in both conditions. In
other words, we assume that mean protein level is proportional to mean mRNA
level, and that the proportionality constant is the same in both conditions and
therefore cancels out in the ratio. This is reasonable since the cells in the
two conditions are identical except for the presence of the transcription
factor, and the model assumes that the transcription factor has no direct effect
on translation.

Fold-change has proven a very convenient observable in past
work~\cite{Garcia2011a, Brewster2014, Razo-Mejia2018, Chure2019}. Part of its
utility in dissecting transcriptional regulation is its ratiometric nature,
which removes many secondary effects that are present when making an absolute
gene expression measurement. Also, by measuring otherwise identical cells with
and without a transcription factor present, any biological noise common to both
conditions can be made to cancel away.

Figure~\ref{fig1:means_cartoons}(B) and (C) depicts an array of mathematicized
cartoons for simple repression using both thermodynamic and kinetic models,
respectively, that have appeared in previous literature. For each cartoon, we
calculate the fold-change in mean gene expression as predicted by that model,
deferring most algebraic details to Appendix~\ref{sec:non_bursty}. What we find
is that for all cartoons the fold-change can be written as a Fermi function

\begin{equation}
 \text{fold-change} = \left( 1 + \exp(-\Delta F_R + \log(\rho))  \right)^{-1},
 \label{eq:deltaFR_eq_noneq_equiv}
\end{equation}
where the effective free energy contains two terms: the parameters $\Delta F_R$,
an effective free energy parametrizing the repressor-DNA interaction, and
$\rho$, a term derived from the level of coarse-graining used to model all
repressor-free states. In other words, the effective free energy of the Fermi
function can be written as the additive effect of the regulation given by the
repressor via $\Delta F_R$, and the kinetic scheme used to describe the steps
that lead to a transcriptional event via $\log(\rho)$ (See
Figure~\ref{fig1:means_cartoons}(D), left panel). This implies that all models
collapse to a single master curve as shown in
Figure~\ref{fig1:means_cartoons}(D). We will offer some intuition for why this
master curve exists and discuss why at the level of the mean expression, we are
unable to discriminate ``right'' from ``wrong'' cartoons given only measurements
of fold-changes in expression.

\subsubsection{Two- and Three-state Thermodynamic Models}

We begin our analysis with models 1 and 2 in
Figure~\ref{fig1:means_cartoons}(B). In each of these models the promoter is
idealized as existing in a set of discrete states; the difference being whether
or not the RNAP bound state is included or not. Gene expression is then assumed
to be proportional to the probability of the promoter being in either the empty
state (model 1) or the RNAP-bound state (model (2)). We direct the reader to
Appendix~\ref{sec:non_bursty} for details on the derivation of the fold-change.
For our purposes here, it suffices to state that the functional form of the
fold-change for model 1 is
\begin{equation}
\text{fold-change}
= \left(1 + \frac{R}{N_{NS}} e^{-\beta\Delta\varepsilon_R}\right)^{-1},
\end{equation}
where $R$ is the number of repressors per cell, $N_{NS}$ is the number of
non-specific binding sites where the repressor can bind, $\Delta\varepsilon_R$
is the repressor-operator binding energy, and $\beta \equiv (k_BT)^{-1}$. This
equation matches the form of the master curve in
Figure~\ref{fig1:means_cartoons}(D) with $\rho=1$ and $\Delta F_R =
\beta\Delta\varepsilon_r - \log (R / N_{NS})$. For model 2 we have a similar
situation. The fold-change takes the form
\begin{eqnarray}
\text{fold-change}
&= \left(
1 + \frac{\frac{R}{N_{NS}} e^{-\beta\Delta\varepsilon_R}}
        {1 + \frac{P}{N_{NS}} e^{-\beta\Delta\varepsilon_P}}
\right)^{-1}
\\
&= (1 + \exp(-\Delta F_R + \log\rho))^{-1},
\end{eqnarray}
where $P$ is the number of RNAP per cell, and $\Delta\varepsilon_P$ is the
RNAP-promoter binding energy. For this model we have $\Delta F_R =
\beta\Delta\varepsilon_R - \log(R/N_{NS})$ and $\rho = 1 +
\frac{P}{N_{NS}}\mathrm{e}^{-\beta\Delta\varepsilon_P}$. Thus far, we see that
the two thermodynamic models, despite making different coarse-graining
commitments, result in the same functional form for the fold-change in mean gene
expression.  We now explore how kinetic models fare when faced with computing
the same observable.

\subsubsection{Kinetic models}
One of the main difference between models shown in
Figure~\ref{fig1:means_cartoons}(C), casted in the language of chemical master
equations, compared with the thermodynamic models discussed in the previous 
section is the probability space over which they are built. Rather than keeping
track only of the microstate of the promoter, and assuming that gene expression
is proportional to the probability of the promoter being in a certain
microstate, chemical master equation models are built on the entire probability
state of both the promoter microstate, and the current mRNA count. Therefore, in
order to compute the fold-change, we must compute the mean mRNA count on each of
the promoter microstates, and add them all together~\cite{Sanchez2013}.

We relegate all details of the derivation to Appendix~\ref{sec:non_bursty}, and
highlight here the general findings for all five kinetic models. As already
shown in Figure~\ref{fig1:means_cartoons}(C) and (D), all the kinetic models
explored can be collapsed onto the master curve. Given that the repressor-bound
state only connects to the rest of the promoter dynamics via its binding and
unbinding rates, $k_R^+$ and $k_R^-$ respectively, all models can effectively be
separated into two categories: a single repressor-bound state, and all other
promoter states with different levels of coarse graining. This structure then
guarantees that, at steady-state, detailed balance between these two groups is
satisfied. What this implies is that the steady-state distribution of each of
the non-repressor states has the same functional form with or without the
repressor, allowing us to write the fold-change as a product of the ratio of the
binding and unbinding rates of the promoter, and the promoter details. This is
\mrm{This is based in one of the comments from reviewer \# 2. It is empirically
true for all models we explored, but I don't know how to prove that it is a
general statement for any unimaginable complex model.}
\begin{eqnarray}
\text{fold-change} &= \left( 1 + \frac{k_R^+}{k_R^-} \rho \right)^{-1},\\
&= (1 + \exp(-\Delta F_R + \log(\rho) ))^{-1},
\end{eqnarray}
where $\Delta F_R \equiv -\log(k_R^+/k_R^-)$, and the functional forms of $\rho$
for each model change as shown in Figure~\ref{fig1:means_cartoons}(C). Another
intuitive way two think about these to terms is as follows: in all kinetic
models shown in Figure~\ref{fig1:means_cartoons}(C) the repressor-bound state
can only be reached from a single repressor-free state. The ratio of these two
states --repressor-bound and adjacent repressor-free state-- must remain the
same for all models, regardless of the details included in other promoter states
if $\Delta F_R$ represents an effective free energy of the repressor binding the
DNA operator. The presence of other states then draws probability density from
the promoter being in either of these two states, making the ratio between the
repressor-bound state and \textit{all} repressor-free states different. The log
difference in this ratio is given by $\log(\rho)$. Since model 1 and model 5 of
Figure~\ref{fig1:means_cartoons}(C) consist of a single repressor-free state,
$\rho$ is then necessarily 1 (See Appendix~\ref{sec:non_bursty} for further
details).

The key outcome of our analysis of the models in
Figure~\ref{fig1:means_cartoons} is the existence of a master curve shown
in Figure~\ref{fig1:means_cartoons}(D) to which the fold-change predictions
of all the models collapse. This master curve is parametrized by only two
effective parameters: $\Delta F_R$, which characterizes the number of repressors
and their binding strength to the DNA, and $\rho$, which characterizes all other
features of the promoter architecture. The key assumption underpinning this
result is that no transcription occurs when a repressor is bound to its
operator. Given this outcome, i.e., the degeneracy of the different models at
the level of fold-change, a mean-based metric such as the fold-change that can
be readily measured experimentally is insufficient to discern between these
different levels of coarse-graining.